### 2024-03-01 HelmetEDADamping.Rmd
Add files via upload

---
title: "EDA-Helmet"
author: "Katie Greene"
date: "2024-02-29"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r}
foam_data <- read.csv(file.choose())

```

```{r}

ggplot(data = foam_data, aes(y = damping)) + geom_point(aes(x=freq), colour = 'red') 
ggplot(data = foam_data, aes(y = damping)) + geom_point(aes(x=amp), colour = 'orange')
ggplot(data = foam_data, aes(y = damping)) + geom_point(aes(x=initialStrain), colour = 'green')
ggplot(data = foam_data, aes(y = damping)) + geom_point(aes(x=initialStress), colour = 'blue')
ggplot(data = foam_data, aes(x=cycle, y = damping , fill = factor(ID))) + geom_point()
```

```{r}
ggplot(data = foam_data, aes(x = porosity, y = damping, fill = factor(chemIndex))) + geom_boxplot()

```

```{r}

ggplot(foam_data, aes(x = factor(chemIndex), y = damping)) + geom_boxplot()
ggplot(foam_data, aes(x = factor(porosity), y = damping)) + geom_boxplot()
```

```{r}
ggplot(foam_data, aes(x = factor(cycle), y = damping, fill = factor(cycle))) + geom_boxplot()
```

```{r}
hist(foam_data$damping)
```

```{r}
ggplot(foam_data, aes(x = factor(ID), y = damping)) + geom_point()
tally(foam_data$damping ~ foam_data$ID)
```

### 2024-03-01 AvaEDA.Rmd
Add files via upload

---
title: "Ava's EDA"
output: html_notebook
---

This is an [R Markdown](http://rmarkdown.rstudio.com) Notebook. When you execute code within the notebook, the results appear beneath the code.

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*.

```{r}
library(ggplot2)
library(patchwork)
library(tidyverse)
```

```{r}
setwd("C:/Users/Ava/OneDrive - Colorado School of Mines/Desktop/Mines/Year4/Semester2/Stat Practicum")
foam_data <- read.csv("foamData.csv")
```

```{r}
ggplot(foam_data, aes(x = cycle, y = stiffness)) +
  geom_point(aes(color = porosity)) +
  xlim(0,10000)

ggplot(foam_data, aes(x = cycle, y = stiffness)) +
  geom_point(aes(color = initialStrain)) +
  xlim(0,10000)

ggplot(foam_data, aes(x = cycle, y = stiffness)) +
  geom_point(aes(color = initialStress)) +
  xlim(0,10000)

ggplot(foam_data, aes(x = cycle, y = stiffness)) +
  geom_point(aes(color = chemIndex)) +
  xlim(0,10000)

ggplot(foam_data, aes(x = cycle, y = stiffness)) +
  geom_point(aes(color = freq)) +
  xlim(0,10000)

ggplot(foam_data, aes(x = cycle, y = stiffness)) +
  geom_point(aes(color = amp)) +
  xlim(0,10000)
```
```{r}
ggplot(foam_data, aes(x = cycle, y = stiffness)) +
  geom_point() +
  xlim(0,10000)
```

```{r}
ggplot(foam_data, aes(x = porosity, y = stiffness)) +
  geom_boxplot()

ggplot(foam_data, aes(x = factor(chemIndex), y = stiffness)) +
  geom_boxplot()

ggplot(data= foam_data, aes(x = porosity, y = stiffness, fill = factor(chemIndex))) +
  geom_boxplot() 
```

```{r}
ggplot(foam_data, aes(x = cycle, y = damping)) +
  geom_point(aes(color = porosity))

ggplot(foam_data, aes(x = cycle, y = stiffness)) +
  geom_point(aes(color = initialStrain))
```

### 2024-04-09 mixedmodel.Rmd
Add files via upload

---
title: "mixed model"
output: html_document
date: "2024-04-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
setwd("C:/Users/Ava/OneDrive - Colorado School of Mines/Desktop/Mines/Year4/Semester2/Stat Practicum")
foam_data <- read.csv("foamData.csv")
```

```{r}
foam_data2 <- na.omit(foam_data)

foam_data3 <- filter(foam_data2$cycle<= 10000)
```

```{r}
head(foam_data2)
```
## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
library(lme4)
```

```{r}
model1 <- lmer(stiffnessFrac ~ freq + amp + chemIndex + porosity + porosity:chemIndex + log(cycle) + log(cycle):freq + log(cycle):amp + log(cycle):chemIndex + log(cycle):porosity + log(cycle):porosity:chemIndex + (1 + cycle|ID), data = foam_data2)
```

### 2024-04-18 dampingmodel.Rmd
Add files via upload

---
title: "mixedmodel2"
output: html_document
date: "2024-04-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(lme4)
library(Matrix)
library(tidyverse)
library(lmerTest)
library(leaps)
```

```{r cars}
setwd("C:/Users/Ava/OneDrive - Colorado School of Mines/Desktop/Mines/Year4/Semester2/Stat Practicum")
foam_data <- read.csv("foamData.csv")
```

```{r}
foam_data_scaled <- foam_data |>
  filter(cycle <= 10000 & cycle > 0) 
```

```{r}
head(foam_data_scaled)
```

```{r}
coolModel <- lmer(stiffnessFrac ~ porosity + chemIndex + freq + amp + porosity:chemIndex + sqrt_cycle + (porosity + chemIndex + freq + amp + porosity:chemIndex):sqrt_cycle  + (1+sqrt_cycle|ID), data=foam_data_scaled)

summary(coolModel)
anova(coolModel)
plot(coolModel)
```

```{r}
coolModel2 <- lmer(stiffnessFrac ~ chemIndex + freq + amp + sqrt_cycle + (chemIndex + freq + amp):sqrt_cycle  + (1+sqrt_cycle|ID), data=foam_data_scaled)

summary(coolModel2)
anova(coolModel2)
plot(coolModel2)
```
```{r}
# forward selection
regfit.fwd <- regsubsets(stiffnessFrac ~ ID + chemIndex + porosity + freq + amp + initialStrain + initialStress + cycle, data = foam_data_scaled, nvmax = 19, method = "forward")

summary(regfit.fwd)
```
```{r}
regfit.fwd <- regsubsets(log(dampingFrac) ~ ID + chemIndex + porosity + freq + amp + initialStrain + initialStress + log(cycle), data = foam_data_scaled, nvmax = 19, method = "forward")

summary(regfit.fwd)
```

```{r}
regfit.full <- regsubsets(log(dampingFrac) ~ chemIndex + porosity + freq + amp + initialStrain + initialStress + log(cycle) + chemIndex:porosity + (chemIndex + porosity + freq + amp + initialStrain + initialStress + chemIndex:porosity)*log(cycle) + (1 + log(cycle)|ID), data = foam_data_scaled)

plot(regfit.full, scale = "r2")
plot(regfit.full, scale = "adjr2")
plot(regfit.full, scale = "Cp")
plot(regfit.full, scale = "bic")

```

```{r}
BICmodelD <- lmer(log(dampingFrac) ~ porosity + chemIndex + freq + log(cycle) + porosity:log(cycle) + freq:log(cycle) + chemIndex:log(cycle) + (1 + log(cycle)|ID), data = foam_data_scaled)

summary(BICmodelD)
```

```{r}
plot(BICmodelD)
qqnorm(resid(BICmodelD))
qqline(resid(BICmodelD))
```

```{r}
regfit.full <- regsubsets(log(stiffnessFrac) ~ chemIndex + porosity + freq + amp + initialStrain + initialStress + log(cycle) + (chemIndex + porosity + freq + amp + initialStrain + initialStress)*log(cycle) + (1 + log(cycle)|ID), data = foam_data_scaled)

plot(regfit.full, scale = "r2")
plot(regfit.full, scale = "adjr2")
plot(regfit.full, scale = "Cp")
plot(regfit.full, scale = "bic")
```

```{r}
BICmodelS <- lmer(log(stiffnessFrac) ~ freq + amp + log(cycle) + freq:log(cycle) + amp:log(cycle) + (1 +log(cycle)|ID), data = foam_data_scaled)

summary(BICmodelD)
```

```{r}
plot(BICmodelS)
qqnorm(resid(BICmodelS))
qqline(resid(BICmodelS))
```

### 2024-04-18 finalModels.Rmd
Add files via upload

---
title: "Final Models"
author: "Yellow Team"
date: "2024-04-18"
output: pdf_document
---

```{r,echo=FALSE,message=FALSE,warning=FALSE}
# Set so that long lines in R will be wrapped:
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80), tidy=TRUE)
```

```{r}
library(lme4)
library(lmerTest)
library(leaps)
library(tidyverse)
```

## Data wrangling

```{r}
foam_data <- read.csv("HelmetFinal.csv")
```

```{r}
foam_data_scaled <- foam_data |>
  filter(cycle <= 10000 & cycle > 0)
```

## Damping Model

```{r}
regfit.full <- regsubsets(log(dampingFrac) ~ chemIndex + porosity + freq + amp + initialStrain + initialStress + log(cycle) + chemIndex:porosity + (chemIndex + porosity + freq + amp + initialStrain + initialStress + chemIndex:porosity)*log(cycle) + (1 + log(cycle)|ID), data = foam_data_scaled)

plot(regfit.full, scale = "bic")

```

**Theoretical Model**

$log(dampingFrac) = \alpha_0 + \alpha_1*I(porosity = low) + \alpha_2*I(chemIndex = 2) + \alpha_3*I(chemIndex = 3) + \alpha_4*freq + \beta_0*log(cycle) + \beta_1*(I(porosity = low)*log(cycle)) + \beta_3*(I(chemIndex = 2)*log(cycle)) + \beta_4*(I(chemIndex = 3)*log(cycle)) + \beta_5*(freq*log(cycle)) + u_i + v_i*log(cycle) + \epsilon_{ij}$

```{r}
dampModel <- lmer(log(dampingFrac) ~ porosity + factor(chemIndex) + freq + log(cycle) + porosity:log(cycle) + freq:log(cycle) + factor(chemIndex):log(cycle) + (1 + log(cycle)|ID), data = foam_data_scaled)

plot(dampModel)
qqnorm(resid(dampModel))
qqline(resid(dampModel))
```

```{r}
summary(dampModel)
```

```{r}
re <- ranef(dampModel)$ID

plot(re)
```

## Stiffness Model

```{r}
regfit.full <- regsubsets(log(stiffnessFrac) ~ chemIndex + porosity + freq + amp + initialStrain + initialStress + log(cycle) + (chemIndex + porosity + freq + amp + initialStrain + initialStress)*log(cycle) + (1 + log(cycle)|ID), data = foam_data_scaled)
#regfit.full <- regsubsets(stiffnessFrac ~ ., data = foam_data_scaled)
plot(regfit.full, scale = 'bic')
```

**Theoretical Model**

$log(stiffnessFrac) = \alpha_0 + \alpha_1*I(porosity = low)) + \alpha_2*I(chemIndex = 2) + \alpha_3*I(chemIndex = 3) + \alpha_4*I(chemIndex = 2)*I(porosity = low) + \alpha_5*I(chemIndex = 3)*I(porosity = low) + \beta_0*log(cycle) +  \beta_1*log(cycle)*I(porosity = low) + \beta_2*log(cycle)*I(chemIndex = 2) + \beta_3*log(cycle)*I(chemIndex = 2)*I(porosity = low) + \beta_4*log(cycle)*I(chemIndex = 3)*I(porosity = low) + u_i + v_i*log(cycle) + \epsilon_{ij}$

```{r}
FactorModel <- lmer(log(stiffnessFrac) ~  porosity + factor(chemIndex) + factor(chemIndex):porosity +  log(cycle) +  factor(chemIndex):log(cycle) + porosity:log(cycle) + factor(chemIndex):porosity:log(cycle) + (1 + log(cycle)|ID), data = foam_data_scaled)
plot(FactorModel)
qqnorm(resid(FactorModel))
qqline(resid(FactorModel))
```

```{r}
summary(FactorModel)
```

```{r}
re <- ranef(FactorModel)$ID

plot(re)
```

### 2024-04-23 draft1.Rmd
Add files via upload

---
title: "report rough draft"
output: html_document
date: "2024-04-18"
---

---
title: "Helmet Foam Case Study Report"
author: 'Katie Greene, Ava Harman, Chris Seery, Ian Stonecypher'
date: 'April, 2024'
format: pdf
---

```{r}
#| echo: false
#| include: false
#| label: front-matter
#| warning: false
# clean up & set default chunk options
rm(list = ls())
knitr::opts_chunk$set(echo = FALSE)

# packages
library(tidyverse) # for example
library(ggplot2) # for example
library(car)       # for example
library(tinytex)   # for example
library(patchwork)
foam_data <- read.csv("foamData.csv")

foam_data_scaled <- foam_data |>
  filter(cycle <= 10000 & cycle > 0)

```

## 1. Project Description

This project comes to us from Moria Foster, from the Lawrence Livermore National Laboratory. Foster has been running experiments on the Colorado School of Mines campus to study the fatigue responses of stiffness and damping in different varieties of foam. In her designed experiment, pieces of foam were put into a universal load frame, which repeatedly compressed and decompressed each piece for thousands of cycles, and each piece had its stiffness and damping measured periodically. These measurements, as well as the parameters of each experimental run, were given to us to aid in the experiment’s analysis stage.

Foster’s objective is to understand what effects the changes in stiffness and damping in a piece of foam under repeated light compressions. The goal of our analysis is to discover patterns and build models to provide this understanding. With this understanding, the U.S. Military could improve the foam padding in marine special combat helmets, ideally so they can survive at least 10 years of regular use.

## 1.1 Research Questions

**Question 1:** What parameters are impacting the stiffness and damping of foam used in military helmets?

**Question 2:** How are these parameters impacting the fatigue response in stiffness and damping over time?

## 1.2 Variables

Explanatory Variables:

|    Variable Name    |                   Description                                  |:--------------------|:---------------------------------------------------------------------------------------------|
        ID            |  the unique identifier of the different foam pieces
     chemIndex        |  the chemical makeup type of the foam. 1 = 79, 2 = 100, 3 = 121
      porosity        |  the porosity level of the foam. Low = 71%, high = 81%
       freq           |  the number of compressions per second (hz)
        amp           |  the amplitude of the compressions
    initialStrain     |  the initial measure of the strain of foam. (Originally labeled “strain”)
    initialStress     |  the initial measure of the stress of the foam. (Originally labeled “stresskpa”)
       cycle          |  the cycle number the measurement was taken at

Aside from cycle, all of these explanatory variables are constant across all measurements of an individual foam piece.

Response Variables:

|    Variable Name    |                   Description                                        |:--------------------|:---------------------------------------------------------------------------------------------|
     stiffness        |  measurement of the foam’s stiffness. (Originally labeled “E1_MPA”)
      damping         |  measurement of the foam’s damping. (Originally labeled “delta_rad”)
    stiffnessFrac     |  stiffness / (first stiffness measurement)
     dampingFrac      |  tan(damping)/ tan(first damping measurement)

The formulas for stiffnessFrac and dampingFrac were specified by Foster. She believed these would be more useful alternatives to studying raw stiffness and damping.

We have initial concerns with the amount of data we have; whether there is enough to make a cause and effect conclusion.

We reformatted our data from wide to long to make each observation represent one measurement as opposed to one piece of foam. This made the data much easier to access for statistical analysis.

The sample code contained the chemical index, the porosity and the sample id of each foam. Using the separate function in dplyr, we separated these characteristics into distinct columns.

We took out samples after 10,000…

#  2. Exploratory Data Analysis (EDA)

*Description of important variables.*
- *summary statistics*
- *missing Data*
- *unusual or concerning observations*

*Relevant summary figures that provide intuition for the research questions and/or inform important modeling decisions (correlation structure, etc).*

*Describe important outcomes of the EDA that the reader should notice.*

# 3. Statistical Analysis

```{r}
#| echo: false
#| label: fig-2
#| warning: false
#| fig-cap: "log(stiffnessFrac) vs log(cycle) by freq, chemIndex, and porosity"
#| fig-align: "center"
#| fig-width: 15
#| fig-height: 5
ggplot(foam_data_scaled, aes(y = stiffnessFrac, x = cycle)) + geom_point()
sp1 <- ggplot(foam_data_scaled, aes(y = log(stiffnessFrac), x = log(cycle))) + geom_point(aes(color = freq))
sp2 <- ggplot(foam_data_scaled, aes(y = log(stiffnessFrac), x = log(cycle))) + geom_point(aes(color = factor(chemIndex)))
sp3 <- ggplot(foam_data_scaled, aes(y = log(stiffnessFrac), x = log(cycle))) + geom_point(aes(color = porosity))

combined <- sp1 + sp2 + sp3 + plot_layout(ncol = 2)

combined
```
```{r}
ggplot(foam_data_scaled, aes(y = dampingFrac, x = cycle)) + geom_point()
ggplot(foam_data_scaled, aes(y = log(dampingFrac), x = log(cycle))) + geom_point(aes(color = freq))
ggplot(foam_data_scaled, aes(y = log(dampingFrac), x = log(cycle))) + geom_point(aes(color = factor(chemIndex)))
ggplot(foam_data_scaled, aes(y = log(dampingFrac), x = log(cycle))) + geom_point(aes(color = porosity))
```
*Describe statistical analysis in terms that your client will be able to understand and perhaps use in presenting their own research.  This section should summarize all relevant analyses that lead to your final conclusions/decisions/recommendations, but should NEVER include visible code (folded code chunks are fine) or unformatted software output.*

- *Model Assumptions & how checked/verified (details left to appendix)*
- *Interpretation of estimates/statistics/p-values IN CONTEXT*
- *Enough detail that the model/analysis could be reproduced by another statistician*

# 4. Recommendations

*Succinct response to each question laid out in 1.2.  This a much shorter version of section 3, and focuses on conclusions rather than the analyses.*

**Question 1:**

**Question 2:**

*(all research questions in 1.1 and/or statistical question in 1.2 should be addressed, in the same order)*

# 5. Resources

*List resources that your client might find useful*

# 6. Additional Considerations

- *Limitations to the recommendations*
- *Concerns you may have about the study; suggestions for similar studies in future*
- *Technical comments*

# Technical Appendix

*Detailed information and a copy of code and or software results.  Additional graphs and supporting figures may also be placed in the appendix.*

### R Script
```{r ref.label=c('front-matter', 'fig-example1','tbl-summarystats', 'fig-example2')}
#| echo: true
#| eval: false

# Reprinted code chunks used previously for analysis
```

<!--
Figure-making Examples:

For example, notice the way @fig-example1 shows a scatterplot.

```{r}
#| echo: false
#| label: fig-example1
#| warning: false
#| fig-cap: "Fuel Efficiency as it relates to vehicle weight for a sample of cars."
#| fig-align: "center"
#| fig-width: 3
#| fig-height: 3

data("mtcars")
gf_point(mpg~ wt, data = mtcars)
```

These are interesting remarks about data summaries.  The overall mean fuel economy in the data is `r mean(mtcars$mpg, na.rm = TRUE)` miles per gallon, which I can calculate in-line since it's very simple.  Most code requires a code chunk, but raw code and output should never appear in the body of the report.  Here's a nice table of summary statistics formatted for the report in @tbl-summarystats.

```{r}
#| echo: false
#| label: tbl-summarystats
#| tbl-cap: "Summary Statistics for Engine Displacement by Cylinders"

favstats(disp ~ cyl, data = mtcars) %>%
  knitr::kable()

```

Look at @fig-example2 to see what figure with subfigures looks like.

```{r}
#| label: fig-example2
#| layout-ncol: 2
#| fig-cap: "A figure with two subfigures!"
#| fig-subcap:  
#|   - "Line Plot 1"
#|   - "Line Plot 2"

plot(cars)
plot(pressure)
```

-->

### 2024-04-24 draft1.Rmd
Update draft1.Rmd

@@
 library(patchwork)
+library(lme4)
 foam_data <- read.csv("foamData.csv")
@@
   filter(cycle <= 10000 & cycle > 0)
+
+FactorModel <- lmer(log(stiffnessFrac) ~  porosity + factor(chemIndex) + factor(chemIndex):porosity +  log(cycle) +  factor(chemIndex):log(cycle) + porosity:log(cycle) + factor(chemIndex):porosity:log(cycle) + (1 + log(cycle)|ID), data = foam_data_scaled)
+
+dampingModel <- lmer(log(dampingFrac) ~ porosity + factor(chemIndex) + freq + log(cycle) + porosity:log(cycle) + freq:log(cycle) + factor(chemIndex):log(cycle) + (1 + log(cycle)|ID), data = foam_data_scaled)
 
@@
 
-*Description of important variables.*
-- *summary statistics*
-- *missing Data*
-- *unusual or concerning observations*
+There was no missing data, but some foam samples had observations taken at very high cycle numbers (above 10000) while observations from other samples stopped at 10000.
+By Foster’s recommendation, all observations after 10000 cycles were ignored.
 
-*Relevant summary figures that provide intuition for the research questions and/or inform important modeling decisions (correlation structure, etc).*
+There were a few concerning observations with negative values of stiffness and/or damping.
+At Foster’s instruction, we converted the negative stiffness values to positive, and left the negative damping values how they were.
 
-*Describe important outcomes of the EDA that the reader should notice.*
+```{r}
+foam_Samples <- foam_data[ ,c("ID", "chemIndex", "porosity", "freq", "amp", "initialStrain", "initialStress")]
+foam_Samples <- unique(foam_Samples)
 
-# 3. Statistical Analysis
+count(foam_Samples, chemIndex)
+count(foam_Samples, porosity)
+count(foam_Samples, chemIndex, porosity)
+```
+
+The counts above show that unequal amounts of each foam type were tested.
+For instance, only five low-porosity foams with chemical index 2 were tested, and there were more foams with chemical index 1 than there were with chemical indexes 2 and 3.
+It will be more difficult to make solid conclusions about types of foam with fewer samples.
+
+```{r}
+#| echo: false
+#| label: fig-1
+#| warning: false
+#| fig-cap: "Histogram of Foam Samples"
+#| fig-align: "center"
+#| fig-width: 10
+#| fig-height: 4
+#| 
+ggplot(foam_Samples, aes(x=freq)) +
+  geom_histogram()
+```
+
+The treatment levels of the foam pieces were uneven as well.
+The histogram above shows that some compression frequencies were tried on many more samples than other frequencies were.
+If one of these underrepresented frequencies has an impact on the response, we will not be able to conclude this confidently.
 
@@
 #| warning: false
+#| fig-cap: "siffnessFrac vs cycle"
+#| fig-align: "center"
+#| fig-width: 10
+#| fig-height: 4
+
+p1 <- ggplot(foam_data_scaled, aes(y = stiffnessFrac, x = cycle)) + geom_point()
+
+p2 <- ggplot(foam_data_scaled, aes(y = dampingFrac, x = cycle)) + geom_point()
+
+combined <- p1 + p2
+
+combined
+```
+
+We first explored the relationship between our responses, stiffnessFrac and dampingFrac, and cycle.
+These have two very interesting trends.
+Then, we decided to log both the response and cycle in order to get a more linear trend that is better for interpretation.
+
+```{r}
+#| echo: false
+#| label: fig-3
+#| warning: false
 #| fig-cap: "log(stiffnessFrac) vs log(cycle) by freq, chemIndex, and porosity"
 #| fig-align: "center"
-#| fig-width: 15
+#| fig-width: 18
 #| fig-height: 5
-ggplot(foam_data_scaled, aes(y = stiffnessFrac, x = cycle)) + geom_point()
+
 sp1 <- ggplot(foam_data_scaled, aes(y = log(stiffnessFrac), x = log(cycle))) + geom_point(aes(color = freq))
@@
 
-combined <- sp1 + sp2 + sp3 + plot_layout(ncol = 2)
+combined <- sp1 + sp2 + sp3
 
@@
 ```
+
+These plots show the results of log(stiffnessFrac) vs log(cycle) separated by frequency, chemIndex, and porosity.
+These show the trends of the three main predictors we expect to use in our model and how they impact one of the responses.
+From the frequency graph, we see that higher frequencies have a more negative trend than lower frequencies.
+The chemIndex graph is not super conclusive, however the chemIndex 1 (actual chemIndex 79) seems to have a more positive trend while chemIndex 3 (actual chemIndex 121) seems to be slightly more negative and chemIndex 2 (actual chemIndex 100) has both positive and negative trends.
+Finally, the porosity graph is not super conclusive because both high and low porosities have both positive and negative slopes.
+
 ```{r}
-ggplot(foam_data_scaled, aes(y = dampingFrac, x = cycle)) + geom_point()
-ggplot(foam_data_scaled, aes(y = log(dampingFrac), x = log(cycle))) + geom_point(aes(color = freq))
-ggplot(foam_data_scaled, aes(y = log(dampingFrac), x = log(cycle))) + geom_point(aes(color = factor(chemIndex)))
-ggplot(foam_data_scaled, aes(y = log(dampingFrac), x = log(cycle))) + geom_point(aes(color = porosity))
+#| echo: false
+#| label: fig-4
+#| warning: false
+#| fig-cap: "log(dampingFrac) vs log(cycle) by freq, chemIndex, and porosity"
+#| fig-align: "center"
+#| fig-width: 18
+#| fig-height: 5
+#| 
+dp1 <- ggplot(foam_data_scaled, aes(y = log(dampingFrac), x = log(cycle))) + geom_point(aes(color = freq))
+dp2 <- ggplot(foam_data_scaled, aes(y = log(dampingFrac), x = log(cycle))) + geom_point(aes(color = factor(chemIndex)))
+dp3 <- ggplot(foam_data_scaled, aes(y = log(dampingFrac), x = log(cycle))) + geom_point(aes(color = porosity))
+
+combined <- dp1 + dp2 + dp3
+
+combined
 ```
-*Describe statistical analysis in terms that your client will be able to understand and perhaps use in presenting their own research.
-This section should summarize all relevant analyses that lead to your final conclusions/decisions/recommendations, but should NEVER include visible code (folded code chunks are fine) or unformatted software output.*
 
-- *Model Assumptions & how checked/verified (details left to appendix)*
-- *Interpretation of estimates/statistics/p-values IN CONTEXT*
-- *Enough detail that the model/analysis could be reproduced by another statistician*
+These plots are the same as the previous plots except with log(dampingFrac) as the response.
+From the frequency graph, we see that lower frequencies have a more negative trend than lower frequencies.
+The chemIndex graph is not super conclusive since all three chemical indexes have both positive and negative trends.
+Finally, the porosity graph is not super conclusive because both high and low porosities have both positive and negative slopes, however the higher porosity tends to be more positive while the lower porosity is more negative.
+
+```{r}
+#| echo: false
+#| label: fig-5
+#| warning: false
+#| fig-cap: "stiffnessFrac vs cycle by chemIndex and porosity"
+#| fig-align: "center"
+#| fig-width: 18
+#| fig-height: 8
+#| 
+sfc1 <- ggplot(foam_data[foam_data$cycle <= 10000 & foam_data$cycle > 0 & foam_data$chemIndex == 1 & foam_data$porosity == "high",], aes(x=cycle, y=stiffnessFrac)) +
+  geom_point(color="red", size=3) + 
+  geom_hline(yintercept = 1, linetype="dashed") +
+  ylim(0.7, 1.15)
+sfc2 <- ggplot(foam_data[foam_data$cycle <= 10000 & foam_data$cycle > 0 & foam_data$chemIndex == 2 & foam_data$porosity == "high",], aes(x=cycle, y=stiffnessFrac)) +
+  geom_point(color="green", size=3) + 
+  geom_hline(yintercept = 1, linetype="dashed") +
+  ylim(0.7, 1.15)
+sfc3 <- ggplot(foam_data[foam_data$cycle <= 10000 & foam_data$cycle > 0 & foam_data$chemIndex == 3 & foam_data$porosity == "high",], aes(x=cycle, y=stiffnessFrac)) +
+  geom_point(color="blue", size=3) + 
+  geom_hline(yintercept = 1, linetype="dashed") +
+  ylim(0.7, 1.15)
+sfc4 <- ggplot(foam_data[foam_data$cycle <= 10000 & foam_data$cycle > 0 & foam_data$chemIndex == 1 & foam_data$porosity == "low",], aes(x=cycle, y=stiffnessFrac)) +
+  geom_point(color="red", size=3) + 
+  geom_hline(yintercept = 1, linetype="dashed") +
+  ylim(0.7, 1.15)
+sfc5 <- ggplot(foam_data[foam_data$cycle <= 10000 & foam_data$cycle > 0 & foam_data$chemIndex == 2 & foam_data$porosity == "low",], aes(x=cycle, y=stiffnessFrac)) +
+  geom_point(color="green", size=3) + 
+  geom_hline(yintercept = 1, linetype="dashed") +
+  ylim(0.7, 1.15)
+sfc6 <- ggplot(foam_data[foam_data$cycle <= 10000 & foam_data$cycle > 0 & foam_data$chemIndex == 3 & foam_data$porosity == "low",], aes(x=cycle, y=stiffnessFrac)) +
+  geom_point(col="blue", size=3) + 
+  geom_hline(yintercept = 1, linetype="dashed") +
+  ylim(0.7, 1.15)
+
+combined <- sfc1 + sfc2 + sfc3 + sfc4 + sfc5 + sfc6 + plot_layout(ncol = 3)
+combined
+```
+
+These plots separate the graphs from above to show a better representation on how the interaction between chemical index and porosity impact stiffnessFrac.
+The top row is high porosity and from left to right the chemical index goes 79, 100, 121.
+
+```{r}
+#| echo: false
+#| label: fig-6
+#| warning: false
+#| fig-cap: "dampingFrac vs cycle for chemIndex and porosity"
+#| fig-align: "center"
+#| fig-width: 18
+#| fig-height: 8
+#| 
+dfc1 <- ggplot(foam_data[foam_data$cycle <= 10000 & foam_data$cycle > 0 & foam_data$chemIndex == 1 & foam_data$porosity == "high",], aes(x=cycle, y=dampingFrac)) +
+  geom_point(color="red", size=3) + 
+  geom_hline(yintercept = 1, linetype="dashed") +
+  ylim(0.7, 1.15)
+dfc2 <- ggplot(foam_data[foam_data$cycle <= 10000 & foam_data$cycle > 0 & foam_data$chemIndex == 2 & foam_data$porosity == "high",], aes(x=cycle, y=dampingFrac)) +
+  geom_point(color="green", size=3) + 
+  geom_hline(yintercept = 1, linetype="dashed") +
+  ylim(0.7, 1.15)
+dfc3 <- ggplot(foam_data[foam_data$cycle <= 10000 & foam_data$cycle > 0 & foam_data$chemIndex == 3 & foam_data$porosity == "high",], aes(x=cycle, y=dampingFrac)) +
+  geom_point(color="blue", size=3) + 
+  geom_hline(yintercept = 1, linetype="dashed") +
+  ylim(0.7, 1.15)
+dfc4 <- ggplot(foam_data[foam_data$cycle <= 10000 & foam_data$cycle > 0 & foam_data$chemIndex == 1 & foam_data$porosity == "low",], aes(x=cycle, y=dampingFrac)) +
+  geom_point(color="red", size=3) + 
+  geom_hline(yintercept = 1, linetype="dashed") +
+  ylim(0.7, 1.15)
+dfc5 <- ggplot(foam_data[foam_data$cycle <= 10000 & foam_data$cycle > 0 & foam_data$chemIndex == 2 & foam_data$porosity == "low",], aes(x=cycle, y=dampingFrac)) +
+  geom_point(color="green", size=3) + 
+  geom_hline(yintercept = 1, linetype="dashed") +
+  ylim(0.7, 1.15)
+dfc6 <- ggplot(foam_data[foam_data$cycle <= 10000 & foam_data$cycle > 0 & foam_data$chemIndex == 3 & foam_data$porosity == "low",], aes(x=cycle, y=dampingFrac)) +
+  geom_point(col="blue", size=3) + 
+  geom_hline(yintercept = 1, linetype="dashed") +
+  ylim(0.7, 1.15)
+
+combined <- dfc1 + dfc2 + dfc3 + dfc4 + dfc5 + dfc6 + plot_layout(ncol = 3)
+combined
+```
+
+These plots are the same as the plots above except with dampingFrac as the repsonse.
+
+# 3. Statistical Analysis
+
+Based on the initial research questions, we will be developing two separate models: one for stiffness fraction and the other for damping fraction.
+Due to the experimental design and the fact that our observations are not independent of each other, we used a mixed model format for our predictions.
+
+In order to select our variables, we first started with a mixed model where all variables aside from cycle were treated as level two variables and log(cycle) was treated as a level one variable.
+In this context, the level two variables on their own will impact the intercept of our model and the interaction between log(cycle) and the other variables will impact the slope of our model.
+After starting with this full model, we manually removed variables from the model to increase the significance of the coefficients and to minimize the REML.
+From this process, we were able to create a model using the following predictors:
+
+|    Stiffness Model Predictors    |        Damping Model Predictors                                  |:---------------------------------|:------------------------------------|
+            porosity               |             porosity
+           chemIndex               |             chemIndex
+           log(cycle)              |               freq
+       porosity:chemIndex          |             log(cycle)
+       chemIndex:log(cycle)        |          porosity:log(cycle)
+       porosity:log(cycle)         |         chemIndex:log(cycle)
+  porosity:chemIndex:log(cycle)    |            freq:log(cycle)
+
+# 3.1 Stiffness
+
+To best predict the value of log(stiffness fraction), we used the following model:
+$$log(stiffnessFrac) = \alpha_0 + \alpha_1*I(porosity = low)) + \alpha_2*I(chemIndex = 2) + \alpha_3*I(chemIndex = 3) + \alpha_4*I(chemIndex = 2)*I(porosity = low) + \alpha_5*I(chemIndex = 3)*I(porosity = low) + \beta_0*log(cycle) +  \beta_1*log(cycle)*I(porosity = low) + \beta_2*log(cycle)*I(chemIndex = 2) + \beta_3*log(cycle)*I(chemIndex = 2)*I(porosity = low) + \beta_4*log(cycle)*I(chemIndex = 3)*I(porosity = low) + u_i + v_i*log(cycle) + \epsilon_{ij}$$
+
+From the statistical summary below we see a high significance in our coefficients.
+The p-value for every term has a significance below five percent, meaning that there is a significant relationship between each of the terms and the response variable.
+The REML is also extremely small which leads us to conclude that this model fits the data very well and has minimal error.
+The coefficients not relating to cycle are impacting the initial stiffness of the foam at time zero.
+The variable that has the most positive impact on the initial stiffness is having a chemical index of 2.
+
+```{r, echo = FALSE}
+summary(FactorModel)
+```
+
+We also assessed the model to see how it met the conditions of a predictive model.
+From the first plot, we see that there appears to be some sinusoidal behavior at the beginning and the middle of the shape has a larger range than the rest of the plot.
+Since the range of the plot is so small, we can conclude that the equal variance condition is being met.
+We also do not see any flaring or nonlinear movement so the linearity condition is also being met.
+The last plot shows some tailing on the ends but nothing too concerning.
+This concludes that all conditions are being met so this model is valid for the purpose of prediction.
+
+```{r, echo = FALSE}
+plot(FactorModel)
+qqnorm(resid(FactorModel))
+qqline(resid(FactorModel))
+```
+
+Lastly, given that we are using a mixed effects model, it is necessary to explore the random effects associated with the model.
+From the plot below, we notice that as the intercept increases, the slope decreases.
+This is a very defined trend and supports our conclusion that a mixed model should be used.
+We would expect to see that each of the foams starts at a similar stiffness fraction and as time increases, there would be deviations based on the other predictors.
+This plot confirms this behavior.
+
+```{r, echo = FALSE}
+re <- ranef(FactorModel)$ID
+
+plot(re)
+```
+
+#	3.2 Damping
+
+To best predict the value of log(damping fraction), we used this model:
+$$log(dampingFrac) = \alpha_0 + \alpha_1*I(porosity = low) + \alpha_2*I(chemIndex = 2) + \alpha_3*I(chemIndex = 3) + \alpha_4*freq + \beta_0*log(cycle) + \beta_1*(I(porosity = low)*log(cycle)) + \beta_3*(I(chemIndex = 2)*log(cycle)) + \beta_4*(I(chemIndex = 3)*log(cycle)) + \beta_5*(freq*log(cycle)) + u_i + v_i*log(cycle) + \epsilon_{ij}$$.
+
+The summary of this model shows a high significance in almost all of our coefficients through small p-values.
+The only value that does not show a significant p-value is frequency, however we chose to keep that in the model because of how significant the interaction between frequency and log(cycle) is.
+This model has multiple interactions: porosity and log(cycle), frequency and log(cycle), and chemical index and log(cycle).
+These interactions show that if all other variables are held constant, there is a combined effect of these interactions on the response variable.
+Since the REML is very negative which shows that the model fits the data and there is minimal error.
+
+```{r, echo = FALSE}
+summary(dampingModel)
+```
+
+From the diagnostic plots, we first look at the residuals vs fitted plot.
+From this plot we can see that there is a mostly linear trend with only a few straggling points throughout.
+The linearity condition is met since there is no major trend to contradict it.
+Also, the equal variance condition is met since there are only a couple of points that have concerningly high residuals.
+In the normal Q-Q plot, there is some tailing at both ends, but nothing to cause extreme concern so the normality condition is met.
+Since all three conditions are met, this model is valid for the purpose of prediction.
+
+```{r, echo = FALSE}
+plot(dampingModel)
+qqnorm(resid(dampingModel))
+qqline(resid(dampingModel))
+```
+
+Lastly, we must look at the random effects of our model.
+Like the stiffness model, as the intercept increases, the slope decreases.
+Although the trend is not as linear as the stiffness model, it still supports our conclusion that a mixed model should be used.
+Again, we would expect to see that each of the foams starts at a similar dampint fraction and as time increases, there would be deviations based on the other predictors.
+This plot confirms this behavior.
+
+```{r, echo = FALSE}
+re <- ranef(dampingModel)$ID
+
+plot(re)
+```
 
@@
 
-**Question 1:**
+**Question 1:** We see that porosity, chemical index, and time have the most significant effects on damping and stiffness. Frequency also has a significant effect but only on damping.
 
-**Question 2:**
+**Question 2:** For stiffness, we see that the individual predictors on their own will increase the initial stiffness, but as time goes on, foam using chemical index 121 will have the greatest decrease in stiffness.
+The individual predictors will impact damping in a similar way but foam with a chemical index of 79 will see the greatest decay in damping.
+Overall, there appear to be significant impacts on the stiffness and damping values depending on chemical index and porosity
 
@@
 
-- *Limitations to the recommendations*
-- *Concerns you may have about the study; suggestions for similar studies in future*
-- *Technical comments*
+One concern we have about our conclusions is that they were made based on relatively few observations.
+Our model for stiffness fraction, for example, consists of six possible trendlines, based on a foam’s porosity and chemical index, and most combinations of porosity and chemical index had fewer than ten foam samples in this experiment.
+We can confidently say that the coefficient estimates in both our models are non-zero, but the estimates could still be off.
+We would need more observations to get a more reliable estimate of the true coefficients.
+Compounding this problem, other treatment levels like amplitude and initial stress were not used evenly in the experiment.
+For example, a few samples were compressed with abnormal amplitude.
+If this had an effect on those samples’ stiffness and damping, it could have a big impact on the coefficient estimates since there weren’t enough other samples to balance it out.
+It could be that amplitude or some other factor has a significant effect on stiffness and damping, but we could not assess that impact confidently because there were not enough observations at each treatment level.
+For this reason, a predictor that should have been included in the model could have been left out.
+
+Another consideration is that our models predict the log of stiffness fraction and the log of damping fraction linearly with the log of cycle number, which is a poor approximation in some cases.
+It is possible that models that are quadratic in log cycle number could have fit the data with slightly better accuracy, but these models would be harder to interpret.
+So our final models prioritize understanding the relationship between the responses and the predictors, possibly at the expense of accuracy.
+
+For further and future studies, we recommend that more foam samples are tested, so that there are more observations to work with.
+The same number of each type of foam should be used, and the treatments such as amplitude and initial stress should be assigned evenly, so each treatment level is fully tested on each type of foam.
+This will allow for more trustworthy models.
 

### 2024-04-24 draft1.Rmd
Update draft1.Rmd
added extended descriptions for plots and data inclusions/wrangling.

@@
----
-title: "report rough draft"
-output: html_document
-date: "2024-04-18"
----
-
 ---
@@
 dampingModel <- lmer(log(dampingFrac) ~ porosity + factor(chemIndex) + freq + log(cycle) + porosity:log(cycle) + freq:log(cycle) + factor(chemIndex):log(cycle) + (1 + log(cycle)|ID), data = foam_data_scaled)
+
+# Convert intervalCode to a factor with specified levels
+foam_data_scaled$intervalCode <- factor(foam_data_scaled$intervalCode, levels = c("A", "B", "C", "D","E", "F", "G", "H","I", "J", "K", "L","M", "N"))
 
@@
       damping         |  measurement of the foam’s damping. (Originally labeled “delta_rad”)
-    stiffnessFrac     |  stiffness / (first stiffness measurement)
-     dampingFrac      |  tan(damping)/ tan(first damping measurement)
+    stiffnessFrac     |  $$\frac{\text{stiffness}}{\text{first stiffness measurement}}$$
+     dampingFrac      |  $$\frac{\text{tan(damping)}}{\text{tan(first damping measurement)}}$$
 
@@
 
-We have initial concerns with the amount of data we have; whether there is enough to make a cause and effect conclusion.
-
-We reformatted our data from wide to long to make each observation represent one measurement as opposed to one piece of foam. This made the data much easier to access for statistical analysis.
-
-The sample code contained the chemical index, the porosity and the sample id of each foam. Using the separate function in dplyr, we separated these characteristics into distinct columns.
-
-We took out samples after 10,000…
-
 #  2. Exploratory Data Analysis (EDA)
 
-There was no missing data, but some foam samples had observations taken at very high cycle numbers (above 10000) while observations from other samples stopped at 10000.
-By Foster’s recommendation, all observations after 10000 cycles were ignored.
+Concerns regarding the data's adequacy for drawing cause-and-effect conclusions prompted us to reformat the data from wide to long, helping our statistical analysis by ensuring each observation represents one measurement.
+Additionally, to maintain consistency across observations, samples taken beyond 10,000 cycles were removed.
+This step ensured that our analysis focuses on the relevant time frame for evaluating foam responses.
+By doing this we are hoping that our EDA gives us helpful insights into how we can optimize foam padding in these military helmets.
 
@@
 
-The treatment levels of the foam pieces were uneven as well.
-The histogram above shows that some compression frequencies were tried on many more samples than other frequencies were.
-If one of these underrepresented frequencies has an impact on the response, we will not be able to conclude this confidently.
+The histogram shows the distribution of compression frequencies among the foam samples.
+The uneven distribution suggests that certain compression frequencies were tested more frequently than others.
+This non-uniformity in sample distribution might impact the reliability of conclusions drawn from the analysis, especially if frequencies with fewer samples have significant effects on the response variables.
 
@@
 
-p2 <- ggplot(foam_data_scaled, aes(y = dampingFrac, x = cycle)) + geom_point()
+p2 <- ggplot(foam_data_scaled, aes(y = dampingFrac, x = cycle)) + 
+  geom_point() +
+  scale_y_continuous(limits = c(0.75, 1))
 
-combined <- p1 + p2
+p3 <- ggplot(foam_data_scaled, aes(y = dampingFrac, x = cycle)) + 
+  geom_point() +
+  scale_y_continuous(limits = c(-30, 1))
+
+combined <- p1 + p2 + p3
 
 combined
+
 ```
+In order to see how stiffnessFrac and dampingFrac change over time for all foams, we plotted a scatter plot for each vs cycle.
+StiffnessFrac seems to spread out in an umbrella shape, showing that as cycle increases the sample values seem to get further apart.
+For dampingFrac, there seems to be a general downward stretching of values as cycle increases, also showing us that as cycle increases sample values seem to spread out.
+However no major relationship appears to be present so futher analysis is required.
 
-We first explored the relationship between our responses, stiffnessFrac and dampingFrac, and cycle.
-These have two very interesting trends.
-Then, we decided to log both the response and cycle in order to get a more linear trend that is better for interpretation.
+We scaled the damping scatter plot so that the scale was from 0.75-1, given that all but 13 observations were in this range.
+The observations outside of this range all occured in the interval code N, which is the last interval code we have in our filtered data, most likely meaning that these foams broke in some way or the experiment failed.
+
+```{r}
+foam_data_interval_N <- foam_data_scaled[foam_data_scaled$intervalcode == "N", ]
+
+ggplot(foam_data_interval_N, aes(y = stiffnessFrac, x = cycle)) + geom_point()
+
+foam_data_scaled$intervalcode == "N"
+
+```
 
@@
 
-These plots show the results of log(stiffnessFrac) vs log(cycle) separated by frequency, chemIndex, and porosity.
+In order to understand how log(stiffnessFrac) behaves over log(cycle) given different characteristics, we have colored the samples in the plots by frequency, chemical index and porosity.
 These show the trends of the three main predictors we expect to use in our model and how they impact one of the responses.
 From the frequency graph, we see that higher frequencies have a more negative trend than lower frequencies.
-The chemIndex graph is not super conclusive, however the chemIndex 1 (actual chemIndex 79) seems to have a more positive trend while chemIndex 3 (actual chemIndex 121) seems to be slightly more negative and chemIndex 2 (actual chemIndex 100) has both positive and negative trends.
-Finally, the porosity graph is not super conclusive because both high and low porosities have both positive and negative slopes.
+The chemIndex graph is not very conclusive, however the chemIndex 1 (actual chemIndex 79) seems to have a more positive trend while chemIndex 3 (actual chemIndex 121) seems to be slightly more negative and chemIndex 2 (actual chemIndex 100) has both positive and negative trends.
+Finally, the porosity graph is also not very conclusive because both high and low porosities have both positive and negative slopes.
 
@@
 The top row is high porosity and from left to right the chemical index goes 79, 100, 121.
+We can see that the chemical index 79 has consitent samples with a stiffnessFrac greater than 1, wheresas 100 and 121 appear to have all values below 1, meaning the stiffness decreases as cycle increases, which is not what we are looking for in a durable foam.
+Therefore this suggets that chemical index is an important variable in the prediction of stiffnessFrac and we should highly consider using it in our model.
 
@@
 These plots are the same as the plots above except with dampingFrac as the repsonse.
+Again, there appears to be a discernable difference in the dampingFrac over cycle by chemical index.
+Our chemical index of 79 appears to have consistent values between 0 and 1, indicating a general decrease in damping as cycle increases.
+For 121 and 100, we have a range of values above and below 1, with foams containig a chemical index of 121 having more regular and evenly spread values as opposed to 100 having a lot of values above 1 and then a group of vaues in the 0.9 range.
+This gives evidence to suggest that dampingFrac is influenced enough by chemical index to warrant including it in our model.
 

### 2024-04-24 draft1.Rmd
Update draft1.Rmd
EDA revisions

@@
 
-p1 <- ggplot(foam_data_scaled, aes(y = stiffnessFrac, x = cycle)) + geom_point()
+p1 <- ggplot(foam_data_scaled, aes(y = stiffnessFrac, x = cycle, col=factor(ID))) +
+  geom_point() + 
+  theme(legend.position="none")
 
-p2 <- ggplot(foam_data_scaled, aes(y = dampingFrac, x = cycle)) + 
+p2 <- ggplot(foam_data_scaled, aes(y = dampingFrac, x = cycle, col=factor(ID))) + 
   geom_point() +
-  scale_y_continuous(limits = c(0.75, 1))
+  scale_y_continuous(limits = c(0.825, 1)) + 
+  theme(legend.position="none")
 
-p3 <- ggplot(foam_data_scaled, aes(y = dampingFrac, x = cycle)) + 
+p3 <- ggplot(foam_data_scaled, aes(y = dampingFrac, x = cycle, col=factor(ID))) + 
   geom_point() +
-  scale_y_continuous(limits = c(-30, 1))
+  scale_y_continuous(limits = c(-20, 1)) +
+  theme(legend.position="none")
 
@@
 In order to see how stiffnessFrac and dampingFrac change over time for all foams, we plotted a scatter plot for each vs cycle.
-StiffnessFrac seems to spread out in an umbrella shape, showing that as cycle increases the sample values seem to get further apart.
-For dampingFrac, there seems to be a general downward stretching of values as cycle increases, also showing us that as cycle increases sample values seem to spread out.
-However no major relationship appears to be present so futher analysis is required.
+Dots of the same color were measured from the same foam sample.
+In the first plot, StiffnessFrac seems to spread out in an umbrella shape, showing that as cycle increases the sample values not only change, but get further apart.
+Cycle number looks like it has a relationship with stiffnessFrac, but that relationship looks different for each foam sample.
 
-We scaled the damping scatter plot so that the scale was from 0.75-1, given that all but 13 observations were in this range.
-The observations outside of this range all occured in the interval code N, which is the last interval code we have in our filtered data, most likely meaning that these foams broke in some way or the experiment failed.
+In the second plot, there seems to be a general downward stretching of dampingFrac values as cycle increases, also showing us that as cycle increases sample values seem to spread out.
+So cycle number looks like it has a negative relationship with dampingFrac, but that relationship still looks different for each foam sample.
+
+The second plot was scaled so that the scale was from 0.75-1, given that all but 13 dampingFrac observations were in this range.
+The observations outside of this range (included in the unscaled third plot) all occurred in the interval code N, which is the last interval code we have in our filtered data, most likely meaning that these foams broke in some way or the experiment failed.
 
 ```{r}
-foam_data_interval_N <- foam_data_scaled[foam_data_scaled$intervalcode == "N", ]
+foam_data_interval_N <- foam_data_scaled[foam_data_scaled$intervalCode == "N", ]
 
@@
 
-foam_data_scaled$intervalcode == "N"
+foam_data_scaled$intervalCode == "N"
 
@@
 #| warning: false
-#| fig-cap: "log(stiffnessFrac) vs log(cycle) by freq, chemIndex, and porosity"
+#| fig-cap: "log(stiffnessFrac) vs log(cycle) by freq, and log(dampingFrac) vs log(cycle) by freq"
 #| fig-align: "center"
-#| fig-width: 18
+#| fig-width: 10
 #| fig-height: 5
@@
 sp1 <- ggplot(foam_data_scaled, aes(y = log(stiffnessFrac), x = log(cycle))) + geom_point(aes(color = freq))
-sp2 <- ggplot(foam_data_scaled, aes(y = log(stiffnessFrac), x = log(cycle))) + geom_point(aes(color = factor(chemIndex)))
-sp3 <- ggplot(foam_data_scaled, aes(y = log(stiffnessFrac), x = log(cycle))) + geom_point(aes(color = porosity))
+#sp2 <- ggplot(foam_data_scaled, aes(y = log(stiffnessFrac), x = log(cycle))) + geom_point(aes(color = factor(chemIndex)))
+#sp3 <- ggplot(foam_data_scaled, aes(y = log(stiffnessFrac), x = log(cycle))) + geom_point(aes(color = porosity))
 
-combined <- sp1 + sp2 + sp3
+#combined <- sp1 + sp2 + sp3
 
+#combined
+
+dp1 <- ggplot(foam_data_scaled, aes(y = log(dampingFrac), x = log(cycle))) + geom_point(aes(color = freq))
+#dp2 <- ggplot(foam_data_scaled, aes(y = log(dampingFrac), x = log(cycle))) + geom_point(aes(color = factor(chemIndex)))
+#dp3 <- ggplot(foam_data_scaled, aes(y = log(dampingFrac), x = log(cycle))) + geom_point(aes(color = porosity))
+
+#combined <- dp1 + dp2 + dp3
+
+#combined
+
+combined <- sp1 + dp1
 combined
@@
 
-In order to understand how log(stiffnessFrac) behaves over log(cycle) given different characteristics, we have colored the samples in the plots by frequency, chemical index and porosity.
-These show the trends of the three main predictors we expect to use in our model and how they impact one of the responses.
-From the frequency graph, we see that higher frequencies have a more negative trend than lower frequencies.
-The chemIndex graph is not very conclusive, however the chemIndex 1 (actual chemIndex 79) seems to have a more positive trend while chemIndex 3 (actual chemIndex 121) seems to be slightly more negative and chemIndex 2 (actual chemIndex 100) has both positive and negative trends.
-Finally, the porosity graph is also not very conclusive because both high and low porosities have both positive and negative slopes.
+In @fig-3 above, we logged the responses, and plotted them against the log of cycle. The relationships between log(cycle) and the logged responses look more linear than the relationships in @fig-2.
+
+In order to understand how log(stiffnessFrac) and log(dampingFrac) behave over log(cycle) given different frequencies, we also colored the samples in the plots above by frequency.
+For log(stiffnessFrac), we see that higher frequencies have a more negative trend than lower frequencies.
+For log(dampingFrac), we see that lower frequencies have a more negative trend than higher frequencies.
+So frequency looks related to both stiffnessFrac and dampingFrac.
 
@@
 #| warning: false
-#| fig-cap: "log(dampingFrac) vs log(cycle) by freq, chemIndex, and porosity"
+#| fig-cap: "log(stiffnessFrac) vs log(cycle) by chemIndex and porosity"
 #| fig-align: "center"
 #| fig-width: 18
-#| fig-height: 5
+#| fig-height: 8
 #| 
-dp1 <- ggplot(foam_data_scaled, aes(y = log(dampingFrac), x = log(cycle))) + geom_point(aes(color = freq))
-dp2 <- ggplot(foam_data_scaled, aes(y = log(dampingFrac), x = log(cycle))) + geom_point(aes(color = factor(chemIndex)))
-dp3 <- ggplot(foam_data_scaled, aes(y = log(dampingFrac), x = log(cycle))) + geom_point(aes(color = porosity))
+sfc1 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 1 & foam_data$porosity == "high",], aes(x=log(cycle), y=log(stiffnessFrac))) +
+  geom_point(color="red", size=3) + 
+  geom_hline(yintercept = 0, linetype="dashed") +
+  ylim(-0.325, 0.15)
+sfc2 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 2 & foam_data$porosity == "high",], aes(x=log(cycle), y=log(stiffnessFrac))) +
+  geom_point(color="green", size=3) + 
+  geom_hline(yintercept = 0, linetype="dashed") +
+  ylim(-0.325, 0.15)
+sfc3 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 3 & foam_data$porosity == "high",], aes(x=log(cycle), y=log(stiffnessFrac))) +
+  geom_point(color="blue", size=3) + 
+  geom_hline(yintercept = 0, linetype="dashed") +
+  ylim(-0.325, 0.15)
+sfc4 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 1 & foam_data$porosity == "low",], aes(x=log(cycle), y=log(stiffnessFrac))) +
+  geom_point(color="red", size=3) + 
+  geom_hline(yintercept = 0, linetype="dashed") +
+  ylim(-0.325, 0.15)
+sfc5 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 2 & foam_data$porosity == "low",], aes(x=log(cycle), y=log(stiffnessFrac))) +
+  geom_point(color="green", size=3) + 
+  geom_hline(yintercept = 0, linetype="dashed") +
+  ylim(-0.325, 0.15)
+sfc6 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 3 & foam_data$porosity == "low",], aes(x=log(cycle), y=log(stiffnessFrac))) +
+  geom_point(col="blue", size=3) + 
+  geom_hline(yintercept = 0, linetype="dashed") +
+  ylim(-0.325, 0.15)
 
-combined <- dp1 + dp2 + dp3
-
+combined <- sfc1 + sfc2 + sfc3 + sfc4 + sfc5 + sfc6 + plot_layout(ncol = 3)
 combined
@@
 
-These plots are the same as the previous plots except with log(dampingFrac) as the response.
-From the frequency graph, we see that lower frequencies have a more negative trend than lower frequencies.
-The chemIndex graph is not super conclusive since all three chemical indexes have both positive and negative trends.
-Finally, the porosity graph is not super conclusive because both high and low porosities have both positive and negative slopes, however the higher porosity tends to be more positive while the lower porosity is more negative.
+These plots separate the log(stiffnessFrac) vs log(cycle) graph from above to show a better representation on how the interaction between chemical index and porosity impact stiffnessFrac.
+The top row shows only high porosity and the bottom row shows only low porosity samples.
+From left to right, the columns show chemical index 79, 100 and 121.
+We can see that the chemical index 79 has many samples where log(stiffnessFrac) trends upwards, whereas chemical indices 100 and 121 mostly trend downwards.
+This indicates a relationship between log(stiffnessFrac) and chemical index.
+Downward trends mean the stiffness decreases as cycle increases, which is not what we are looking for in a durable foam.
+
+We also see an effect from porosity.
+For chemical index 79, low porosity foam trends downward more often than high porosity foam, and for chemical index 100, low porosity foam trends upward more often than high porosity foam.
+So not only does porosity affect log(stiffnessFrac), but it effects it different ways for different porosities.
+This suggests a porosity:chemIndex interraction term would improve our model.
 
@@
 #| warning: false
-#| fig-cap: "stiffnessFrac vs cycle by chemIndex and porosity"
+#| fig-cap: "log(dampingFrac) vs log(cycle) for chemIndex and porosity"
 #| fig-align: "center"
@@
 #| 
-sfc1 <- ggplot(foam_data[foam_data$cycle <= 10000 & foam_data$cycle > 0 & foam_data$chemIndex == 1 & foam_data$porosity == "high",], aes(x=cycle, y=stiffnessFrac)) +
+dfc1 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 1 & foam_data_scaled$porosity == "high",], aes(x=log(cycle), y=log(dampingFrac))) +
   geom_point(color="red", size=3) + 
-  geom_hline(yintercept = 1, linetype="dashed") +
-  ylim(0.7, 1.15)
-sfc2 <- ggplot(foam_data[foam_data$cycle <= 10000 & foam_data$cycle > 0 & foam_data$chemIndex == 2 & foam_data$porosity == "high",], aes(x=cycle, y=stiffnessFrac)) +
+  geom_hline(yintercept = 0, linetype="dashed") +
+  ylim(-0.175, 0.125)
+dfc2 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 2 & foam_data_scaled$porosity == "high",], aes(x=log(cycle), y=log(dampingFrac))) +
   geom_point(color="green", size=3) + 
-  geom_hline(yintercept = 1, linetype="dashed") +
-  ylim(0.7, 1.15)
-sfc3 <- ggplot(foam_data[foam_data$cycle <= 10000 & foam_data$cycle > 0 & foam_data$chemIndex == 3 & foam_data$porosity == "high",], aes(x=cycle, y=stiffnessFrac)) +
+  geom_hline(yintercept = 0, linetype="dashed") +
+  ylim(-0.175, 0.125)
+dfc3 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 3 & foam_data_scaled$porosity == "high",], aes(x=log(cycle), y=log(dampingFrac))) +
   geom_point(color="blue", size=3) + 
-  geom_hline(yintercept = 1, linetype="dashed") +
-  ylim(0.7, 1.15)
-sfc4 <- ggplot(foam_data[foam_data$cycle <= 10000 & foam_data$cycle > 0 & foam_data$chemIndex == 1 & foam_data$porosity == "low",], aes(x=cycle, y=stiffnessFrac)) +
+  geom_hline(yintercept = 0, linetype="dashed") +
+  ylim(-0.175, 0.125)
+dfc4 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 1 & foam_data_scaled$porosity == "low",], aes(x=log(cycle), y=log(dampingFrac))) +
   geom_point(color="red", size=3) + 
-  geom_hline(yintercept = 1, linetype="dashed") +
-  ylim(0.7, 1.15)
-sfc5 <- ggplot(foam_data[foam_data$cycle <= 10000 & foam_data$cycle > 0 & foam_data$chemIndex == 2 & foam_data$porosity == "low",], aes(x=cycle, y=stiffnessFrac)) +
+  geom_hline(yintercept = 0, linetype="dashed") +
+  ylim(-0.175, 0.125)
+dfc5 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 2 & foam_data_scaled$porosity == "low",], aes(x=log(cycle), y=log(dampingFrac))) +
   geom_point(color="green", size=3) + 
-  geom_hline(yintercept = 1, linetype="dashed") +
-  ylim(0.7, 1.15)
-sfc6 <- ggplot(foam_data[foam_data$cycle <= 10000 & foam_data$cycle > 0 & foam_data$chemIndex == 3 & foam_data$porosity == "low",], aes(x=cycle, y=stiffnessFrac)) +
+  geom_hline(yintercept = 0, linetype="dashed") +
+  ylim(-0.175, 0.125)
+dfc6 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 3 & foam_data_scaled$porosity == "low",], aes(x=log(cycle), y=log(dampingFrac))) +
   geom_point(col="blue", size=3) + 
-  geom_hline(yintercept = 1, linetype="dashed") +
-  ylim(0.7, 1.15)
-
-combined <- sfc1 + sfc2 + sfc3 + sfc4 + sfc5 + sfc6 + plot_layout(ncol = 3)
-combined
-```
-
-These plots separate the graphs from above to show a better representation on how the interaction between chemical index and porosity impact stiffnessFrac.
-The top row is high porosity and from left to right the chemical index goes 79, 100, 121.
-We can see that the chemical index 79 has consitent samples with a stiffnessFrac greater than 1, wheresas 100 and 121 appear to have all values below 1, meaning the stiffness decreases as cycle increases, which is not what we are looking for in a durable foam.
-Therefore this suggets that chemical index is an important variable in the prediction of stiffnessFrac and we should highly consider using it in our model.
-
-```{r}
-#| echo: false
-#| label: fig-6
-#| warning: false
-#| fig-cap: "dampingFrac vs cycle for chemIndex and porosity"
-#| fig-align: "center"
-#| fig-width: 18
-#| fig-height: 8
-#| 
-dfc1 <- ggplot(foam_data[foam_data$cycle <= 10000 & foam_data$cycle > 0 & foam_data$chemIndex == 1 & foam_data$porosity == "high",], aes(x=cycle, y=dampingFrac)) +
-  geom_point(color="red", size=3) + 
-  geom_hline(yintercept = 1, linetype="dashed") +
-  ylim(0.7, 1.15)
-dfc2 <- ggplot(foam_data[foam_data$cycle <= 10000 & foam_data$cycle > 0 & foam_data$chemIndex == 2 & foam_data$porosity == "high",], aes(x=cycle, y=dampingFrac)) +
-  geom_point(color="green", size=3) + 
-  geom_hline(yintercept = 1, linetype="dashed") +
-  ylim(0.7, 1.15)
-dfc3 <- ggplot(foam_data[foam_data$cycle <= 10000 & foam_data$cycle > 0 & foam_data$chemIndex == 3 & foam_data$porosity == "high",], aes(x=cycle, y=dampingFrac)) +
-  geom_point(color="blue", size=3) + 
-  geom_hline(yintercept = 1, linetype="dashed") +
-  ylim(0.7, 1.15)
-dfc4 <- ggplot(foam_data[foam_data$cycle <= 10000 & foam_data$cycle > 0 & foam_data$chemIndex == 1 & foam_data$porosity == "low",], aes(x=cycle, y=dampingFrac)) +
-  geom_point(color="red", size=3) + 
-  geom_hline(yintercept = 1, linetype="dashed") +
-  ylim(0.7, 1.15)
-dfc5 <- ggplot(foam_data[foam_data$cycle <= 10000 & foam_data$cycle > 0 & foam_data$chemIndex == 2 & foam_data$porosity == "low",], aes(x=cycle, y=dampingFrac)) +
-  geom_point(color="green", size=3) + 
-  geom_hline(yintercept = 1, linetype="dashed") +
-  ylim(0.7, 1.15)
-dfc6 <- ggplot(foam_data[foam_data$cycle <= 10000 & foam_data$cycle > 0 & foam_data$chemIndex == 3 & foam_data$porosity == "low",], aes(x=cycle, y=dampingFrac)) +
-  geom_point(col="blue", size=3) + 
-  geom_hline(yintercept = 1, linetype="dashed") +
-  ylim(0.7, 1.15)
+  geom_hline(yintercept = 0, linetype="dashed") +
+  ylim(-0.175, 0.125)
 
@@
 
-These plots are the same as the plots above except with dampingFrac as the repsonse.
-Again, there appears to be a discernable difference in the dampingFrac over cycle by chemical index.
-Our chemical index of 79 appears to have consistent values between 0 and 1, indicating a general decrease in damping as cycle increases.
-For 121 and 100, we have a range of values above and below 1, with foams containig a chemical index of 121 having more regular and evenly spread values as opposed to 100 having a lot of values above 1 and then a group of vaues in the 0.9 range.
-This gives evidence to suggest that dampingFrac is influenced enough by chemical index to warrant including it in our model.
+These plots separate the log(dampingFrac) vs log(cycle) graph from above to show a better representation on how the interaction between chemical index and porosity impact dampingFrac.
+The data shown in the rows and columns are the same as in @fig-4.
+We can see differences in the log(dampingFrac) trends between different chemical indices, so there is likely a relationship present, but this relationship looks less pronounced than it was for the log(stiffnessFrac) trends.
+
+Porosity also appears to have some effect.
+Trends in log(dampingFrac) for chemical index 121 are more similar to each other for low porosity than for high porosity, and trends in log(dampingFrac) for chemical index 79 are bit more similar to each other for high porosity than for low porosity.
+This suggests a porosity:chemIndex interaction term would improve our damping model, but this is still less clear than it was for stiffness.
 

### 2024-04-24 draft1.Rmd
Update draft1.Rmd

@@
+---
+title: "Untitled"
+output: html_document
+date: "2024-04-24"
+---
+
 ---
@@
 library(lme4)
+library(lmerTest)
+library(leaps)
 foam_data <- read.csv("foamData.csv")
@@
 
-|    Stiffness Model Predictors    |        Damping Model Predictors                                  |:---------------------------------|:------------------------------------|
+|    Stiffness Model Predictors    |        Damping Model Predictors                     
+|:---------------------------------|:------------------------------------|
             porosity               |             porosity
@@
 
-*(all research questions in 1.1 and/or statistical question in 1.2 should be addressed, in the same order)*
-
-# 5. Resources
-
-*List resources that your client might find useful*
-
-# 6. Additional Considerations
+# 5. Additional Considerations
 
@@
 #| eval: false
+foam_data <- read.csv("foamData.csv")
 
-# Reprinted code chunks used previously for analysis
+foam_data_scaled <- foam_data |>
+  filter(cycle <= 10000 & cycle > 0)
+
+FactorModel <- lmer(log(stiffnessFrac) ~  porosity + factor(chemIndex) + factor(chemIndex):porosity +  log(cycle) +  factor(chemIndex):log(cycle) + porosity:log(cycle) + factor(chemIndex):porosity:log(cycle) + (1 + log(cycle)|ID), data = foam_data_scaled)
+
+dampingModel <- lmer(log(dampingFrac) ~ porosity + factor(chemIndex) + freq + log(cycle) + porosity:log(cycle) + freq:log(cycle) + factor(chemIndex):log(cycle) + (1 + log(cycle)|ID), data = foam_data_scaled)
+
+# Convert intervalCode to a factor with specified levels
+foam_data_scaled$intervalCode <- factor(foam_data_scaled$intervalCode, levels = c("A", "B", "C", "D","E", "F", "G", "H","I", "J", "K", "L","M", "N"))
+
+# summary of stiffnessFrac model 
+summary(FactorModel)
+
+# residuals and QQ plot of stiffnessFrac model
+plot(FactorModel)
+qqnorm(resid(FactorModel))
+qqline(resid(FactorModel))
+
+# random effects of stiffnessFrac model
+re <- ranef(FactorModel)$ID
+
+plot(re)
+
+# summary of dampingFrac model
+summary(dampingModel)
+
+# residuals and QQ plot of dampingFrac model
+plot(dampingModel)
+qqnorm(resid(dampingModel))
+qqline(resid(dampingModel))
+
+# random effects of dampingFrac model
+re <- ranef(dampingModel)$ID
+
+plot(re)
 ```
-
-<!--
-Figure-making Examples:
-
-For example, notice the way @fig-example1 shows a scatterplot.
 
@@
 #| warning: false
-#| fig-cap: "Fuel Efficiency as it relates to vehicle weight for a sample of cars."
+#| fig-cap: "Backward Selection for stiffnessFrac Predictors"
 #| fig-align: "center"
-#| fig-width: 3
-#| fig-height: 3
-
-data("mtcars")
-gf_point(mpg~ wt, data = mtcars)
+#| fig-width: 7
+#| fig-height: 5
+regfit.full <- regsubsets(log(stiffnessFrac) ~ chemIndex + porosity + freq + amp + initialStrain + initialStress + log(cycle) + (chemIndex + porosity + freq + amp + initialStrain + initialStress)*log(cycle) + (1 + log(cycle)|ID), data = foam_data_scaled)
+#regfit.full <- regsubsets(stiffnessFrac ~ ., data = foam_data_scaled)
+plot(regfit.full, scale = 'bic')
 ```
-
-These are interesting remarks about data summaries.
-The overall mean fuel economy in the data is `r mean(mtcars$mpg, na.rm = TRUE)` miles per gallon, which I can calculate in-line since it's very simple.
-Most code requires a code chunk, but raw code and output should never appear in the body of the report.
-Here's a nice table of summary statistics formatted for the report in @tbl-summarystats.
 
@@
 #| echo: false
-#| label: tbl-summarystats
-#| tbl-cap: "Summary Statistics for Engine Displacement by Cylinders"
+#| label: fig-example2
+#| warning: false
+#| fig-cap: "Backward Selection for dampingFrac Predictors"
+#| fig-align: "center"
+#| fig-width: 7
+#| fig-height: 5
+#| 
+regfit.full <- regsubsets(log(dampingFrac) ~ chemIndex + porosity + freq + amp + initialStrain + initialStress + log(cycle) + chemIndex:porosity + (chemIndex + porosity + freq + amp + initialStrain + initialStress + chemIndex:porosity)*log(cycle) + (1 + log(cycle)|ID), data = foam_data_scaled)
 
-favstats(disp ~ cyl, data = mtcars) %>%
-  knitr::kable()
-
+plot(regfit.full, scale = "bic")
 ```
 
-Look at @fig-example2 to see what figure with subfigures looks like.
-
-```{r}
-#| label: fig-example2
-#| layout-ncol: 2
-#| fig-cap: "A figure with two subfigures!"
-#| fig-subcap:  
-#|   - "Line Plot 1"
-#|   - "Line Plot 2"
-
-plot(cars)
-plot(pressure)
-```
-
--->
+@fig-example1 and @fig-example2 show the plots that we used to pick our predictors for our model.
+Since our main concerns for this analysis were with chemIndex and porosity, we chose the predictors with the highest bic starting at where those two predictors became significant.
+With stiffnessFrac this was at bic = -640 and with dampingFrac this was at -520.

### 2024-04-25 draft1.Rmd
Update draft1.Rmd
Tweaks

@@
 
-Foster’s objective is to understand what effects the changes in stiffness and damping in a piece of foam under repeated light compressions.
+Foster’s objective is to understand what affects the changes in stiffness and damping in a piece of foam under repeated light compressions.
 The goal of our analysis is to discover patterns and build models to provide this understanding.
@@
 
-|    Variable Name    |                   Description                                  |:--------------------|:---------------------------------------------------------------------------------------------|
-        ID            |  the unique identifier of the different foam pieces
-     chemIndex        |  the chemical makeup type of the foam. 1 = 79, 2 = 100, 3 = 121
-      porosity        |  the porosity level of the foam. Low = 71%, high = 81%
-       freq           |  the number of compressions per second (hz)
-        amp           |  the amplitude of the compressions
-    initialStrain     |  the initial measure of the strain of foam. (Originally labeled “strain”)
-    initialStress     |  the initial measure of the stress of the foam. (Originally labeled “stresskpa”)
-       cycle          |  the cycle number the measurement was taken at
+|    Variable Name    |                   Description                                 |
+|:--------------------|:---------------------------------------------------------------------------------------------|
+|        ID            |  the unique identifier of the different foam pieces|
+|     chemIndex        |  the chemical makeup type of the foam. 1 = 79, 2 = 100, 3 = 121|
+|      porosity        |  the porosity level of the foam. Low = 71%, high = 81%|
+|       freq           |  the number of compressions per second (hz)|
+|        amp           |  the amplitude of the compressions|
+|    initialStrain     |  the initial measure of the strain of foam. (Originally labeled “strain”)|
+|    initialStress     |  the initial measure of the stress of the foam. (Originally labeled “stresskpa”)|
+|       cycle          |  the cycle number the measurement was taken at|
 
@@
 
-|    Variable Name    |                   Description                                        |:--------------------|:---------------------------------------------------------------------------------------------|
-     stiffness        |  measurement of the foam’s stiffness. (Originally labeled “E1_MPA”)
-      damping         |  measurement of the foam’s damping. (Originally labeled “delta_rad”)
-    stiffnessFrac     |  $$\frac{\text{stiffness}}{\text{first stiffness measurement}}$$
-     dampingFrac      |  $$\frac{\text{tan(damping)}}{\text{tan(first damping measurement)}}$$
+|    Variable Name    |                   Description                                       |
+|:--------------------|:---------------------------------------------------------------------------------------------|
+|     stiffness        |  measurement of the foam’s stiffness. (Originally labeled “E1_MPA”)|
+|      damping         |  measurement of the foam’s damping. (Originally labeled “delta_rad”)|
+|    stiffnessFrac     |  stiffness / first stiffness measurement|
+|     dampingFrac      |  tan(damping) / tan(first damping measurement)
 
@@
 
-foam_data_scaled$intervalCode == "N"
+#foam_data_scaled$intervalCode == "N"
 

### 2024-04-25 draft1.Rmd
Update draft1.Rmd
More tweaks

@@
 
+Printed below are how many of each type of foam sample was tested:
+
 ```{r}
@@
 The counts above show that unequal amounts of each foam type were tested.
-For instance, only five low-porosity foams with chemical index 2 were tested, and there were more foams with chemical index 1 than there were with chemical indexes 2 and 3.
-It will be more difficult to make solid conclusions about types of foam with fewer samples.
+If a specific type of foam (like low-porosity foam with chemical index 2) has very few observations, it will be more difficult to make solid conclusions about this type of foam.
+Even the foam types with more samples still have relatively few, so making solid conclusions for these could be difficult as well.
 
@@
 #| warning: false
-#| fig-cap: "Histogram of Foam Samples"
+#| fig-cap: "Histogram of Frequencies Tested"
 #| fig-align: "center"
@@
 
-The histogram shows the distribution of compression frequencies among the foam samples.
+The histogram in @fig-1 shows the distribution of compression frequencies among the foam samples.
 The uneven distribution suggests that certain compression frequencies were tested more frequently than others.
@@
 In order to see how stiffnessFrac and dampingFrac change over time for all foams, we plotted a scatter plot for each vs cycle.
-Dots of the same color were measured from the same foam sample.
-In the first plot, StiffnessFrac seems to spread out in an umbrella shape, showing that as cycle increases the sample values not only change, but get further apart.
-Cycle number looks like it has a relationship with stiffnessFrac, but that relationship looks different for each foam sample.
+The dot colors in @fig-2 represent which foam sample they were measured from.
+In the first plot, stiffnessFrac seems to spread out in an umbrella shape, showing that as cycle increases the sample values not only change, but get further apart.
+So cycle number has a relationship with stiffnessFrac, but that relationship looks different for each foam sample.
 
@@
 ```{r}
-foam_data_interval_N <- foam_data_scaled[foam_data_scaled$intervalCode == "N", ]
+#foam_data_interval_N <- foam_data_scaled[foam_data_scaled$intervalCode == "N", ]
 
-ggplot(foam_data_interval_N, aes(y = stiffnessFrac, x = cycle)) + geom_point()
+#ggplot(foam_data_interval_N, aes(y = stiffnessFrac, x = cycle)) + geom_point()
 
@@
 
-In @fig-3 above, we logged the responses, and plotted them against the log of cycle. The relationships between log(cycle) and the logged responses look more linear than the relationships in @fig-2.
+In @fig-3, we logged the responses, and plotted them against the log of cycle. The relationships between log(cycle) and the logged responses look more linear than the relationships in @fig-2.
 
-In order to understand how log(stiffnessFrac) and log(dampingFrac) behave over log(cycle) given different frequencies, we also colored the samples in the plots above by frequency.
+In order to understand how log(stiffnessFrac) and log(dampingFrac) behave over log(cycle) given different frequencies, we also colored the samples in @fig-3 by frequency.
 For log(stiffnessFrac), we see that higher frequencies have a more negative trend than lower frequencies.
@@
 sfc3 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 3 & foam_data$porosity == "high",], aes(x=log(cycle), y=log(stiffnessFrac))) +
-  geom_point(color="blue", size=3) + 
+  geom_point(color="cyan2", size=3) + 
   geom_hline(yintercept = 0, linetype="dashed") +
@@
 sfc4 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 1 & foam_data$porosity == "low",], aes(x=log(cycle), y=log(stiffnessFrac))) +
-  geom_point(color="red", size=3) + 
+  geom_point(color="red3", size=3) + 
   geom_hline(yintercept = 0, linetype="dashed") +
@@
 sfc5 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 2 & foam_data$porosity == "low",], aes(x=log(cycle), y=log(stiffnessFrac))) +
-  geom_point(color="green", size=3) + 
+  geom_point(color="green4", size=3) + 
   geom_hline(yintercept = 0, linetype="dashed") +
@@
 
-These plots separate the log(stiffnessFrac) vs log(cycle) graph from above to show a better representation on how the interaction between chemical index and porosity impact stiffnessFrac.
+The plots in @fig-4 separate the log(stiffnessFrac) vs log(cycle) graph from above to show how the interaction between chemical index and porosity impact log(stiffnessFrac).
 The top row shows only high porosity and the bottom row shows only low porosity samples.
@@
 So not only does porosity affect log(stiffnessFrac), but it effects it different ways for different porosities.
-This suggests a porosity:chemIndex interraction term would improve our model.
+This suggests that a porosity:chemIndex interaction term would improve our model.
 
@@
 dfc3 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 3 & foam_data_scaled$porosity == "high",], aes(x=log(cycle), y=log(dampingFrac))) +
-  geom_point(color="blue", size=3) + 
+  geom_point(color="cyan2", size=3) + 
   geom_hline(yintercept = 0, linetype="dashed") +
@@
 dfc4 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 1 & foam_data_scaled$porosity == "low",], aes(x=log(cycle), y=log(dampingFrac))) +
-  geom_point(color="red", size=3) + 
+  geom_point(color="red3", size=3) + 
   geom_hline(yintercept = 0, linetype="dashed") +
@@
 dfc5 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 2 & foam_data_scaled$porosity == "low",], aes(x=log(cycle), y=log(dampingFrac))) +
-  geom_point(color="green", size=3) + 
+  geom_point(color="green4", size=3) + 
   geom_hline(yintercept = 0, linetype="dashed") +
@@
 
-These plots separate the log(dampingFrac) vs log(cycle) graph from above to show a better representation on how the interaction between chemical index and porosity impact dampingFrac.
+The plots in @fig-5 separate the log(dampingFrac) vs log(cycle) graph from above to show how the interaction between chemical index and porosity impact log(dampingFrac).
 The data shown in the rows and columns are the same as in @fig-4.

### 2024-04-25 draft1.Rmd
Update draft1.Rmd
Even more tweaks

@@
 
-In order to select our variables, we first started with a mixed model where all variables aside from cycle were treated as level two variables and log(cycle) was treated as a level one variable.
+In order to select our variables, we first started with a mixed model where all variables aside from log(cycle) were treated as level two variables and log(cycle) was treated as a level one variable.
 In this context, the level two variables on their own will impact the intercept of our model and the interaction between log(cycle) and the other variables will impact the slope of our model.
@@
 To best predict the value of log(stiffness fraction), we used the following model:
-$$log(stiffnessFrac) = \alpha_0 + \alpha_1*I(porosity = low)) + \alpha_2*I(chemIndex = 2) + \alpha_3*I(chemIndex = 3) + \alpha_4*I(chemIndex = 2)*I(porosity = low) + \alpha_5*I(chemIndex = 3)*I(porosity = low) + \beta_0*log(cycle) +  \beta_1*log(cycle)*I(porosity = low) + \beta_2*log(cycle)*I(chemIndex = 2) + \beta_3*log(cycle)*I(chemIndex = 2)*I(porosity = low) + \beta_4*log(cycle)*I(chemIndex = 3)*I(porosity = low) + u_i + v_i*log(cycle) + \epsilon_{ij}$$
+
+$\log(\text{stiffnessFrac}) = \alpha_0 + \alpha_1*I(\text{porosity} = \text{low})) + \alpha_2*I(\text{chemIndex} = 2)$\newline
+$+\alpha_3*I(\text{chemIndex} = 3)+\alpha_4*I(\text{chemIndex} = 2)*I(\text{porosity} = low)$\newline
+$+\alpha_5*I(\text{chemIndex} = 3)*I(\text{porosity} = low)+\beta_0*\log(\text{cycle}) +  \beta_1*\log(\text{cycle})*I(\text{porosity} = \text{low})$\newline
+$+\beta_2*\log(\text{cycle})*I(\text{chemIndex} = 2)+\beta_3*\log(\text{cycle})*I(\text{chemIndex} = 2)*I(\text{porosity} = \text{low})$\newline
+$+\beta_4*\log(\text{cycle})*I(\text{chemIndex} = 3)*I(\text{porosity} = \text{low})+ u_i + v_i*\log(\text{cycle}) + \epsilon_{ij}$
 
@@
 The p-value for every term has a significance below five percent, meaning that there is a significant relationship between each of the terms and the response variable.
-The REML is also extremely small which leads us to conclude that this model fits the data very well and has minimal error.
+The REML is also extremely low which leads us to conclude that this model fits the data very well and has minimal error.
 The coefficients not relating to cycle are impacting the initial stiffness of the foam at time zero.
@@
 We also assessed the model to see how it met the conditions of a predictive model.
-From the first plot, we see that there appears to be some sinusoidal behavior at the beginning and the middle of the shape has a larger range than the rest of the plot.
+From @fig-6, we see that there appears to be some sinusoidal behavior at the beginning and the middle of the shape has a larger range than the rest of the plot.
 Since the range of the plot is so small, we can conclude that the equal variance condition is being met.
 We also do not see any flaring or nonlinear movement so the linearity condition is also being met.
-The last plot shows some tailing on the ends but nothing too concerning.
+The QQ plot in #fig-7 shows some tailing on the ends but nothing too concerning.
 This concludes that all conditions are being met so this model is valid for the purpose of prediction.
@@
 ```{r, echo = FALSE}
+#| echo: false
+#| label: fig-6
+#| warning: false
+#| fig-cap: "Residual Plot of Stiffness Fraction Model"
+#| fig-align: "center"
+#| fig-width: 7
+#| fig-height: 4
+
 plot(FactorModel)
+```
+
+```{r, echo = FALSE}
+#| echo: false
+#| label: fig-7
+#| warning: false
+#| fig-cap: "QQ Plot of Stiffness Fraction Model"
+#| fig-align: "center"
+#| fig-width: 7
+#| fig-height: 4
+
 qqnorm(resid(FactorModel))
@@
 Lastly, given that we are using a mixed effects model, it is necessary to explore the random effects associated with the model.
-From the plot below, we notice that as the intercept increases, the slope decreases.
+From @fig-8, we notice that as the intercept increases, the slope decreases.
 This is a very defined trend and supports our conclusion that a mixed model should be used.
@@
 ```{r, echo = FALSE}
+#| echo: false
+#| label: fig-8
+#| warning: false
+#| fig-cap: "Intercept Rand Effects vs. Slope Rand Effects"
+#| fig-align: "center"
+#| fig-width: 7
+#| fig-height: 4
+
 re <- ranef(FactorModel)$ID
@@
 To best predict the value of log(damping fraction), we used this model:
-$$log(dampingFrac) = \alpha_0 + \alpha_1*I(porosity = low) + \alpha_2*I(chemIndex = 2) + \alpha_3*I(chemIndex = 3) + \alpha_4*freq + \beta_0*log(cycle) + \beta_1*(I(porosity = low)*log(cycle)) + \beta_3*(I(chemIndex = 2)*log(cycle)) + \beta_4*(I(chemIndex = 3)*log(cycle)) + \beta_5*(freq*log(cycle)) + u_i + v_i*log(cycle) + \epsilon_{ij}$$.
+
+$\log(\text{dampingFrac}) = \alpha_0 + \alpha_1*I(\text{porosity} = \text{low}) + \alpha_2*I(\text{chemIndex} = 2) + \alpha_3*I(\text{chemIndex} = 3)$\newline
+$+\alpha_4*\text{freq} + \beta_0*\log(\text{cycle}) + \beta_1*(I(\text{porosity} = \text{low})*\log(\text{cycle}))$\newline
+$+\beta_3*(I(\text{chemIndex} = 2)*\log(\text{cycle})) + \beta_4*(I(\text{chemIndex} = 3)*\log(\text{cycle}))$\newline
+$+\beta_5*(\text{freq}*\log(\text{cycle})) + u_i + v_i*\log(\text{cycle}) + \epsilon_{ij}$.
 
@@
 These interactions show that if all other variables are held constant, there is a combined effect of these interactions on the response variable.
-Since the REML is very negative which shows that the model fits the data and there is minimal error.
+The REML is very negative, which shows that the model fits the data and there is minimal error.
 
@@
 
-From the diagnostic plots, we first look at the residuals vs fitted plot.
+From the diagnostic plots, we first look at the residuals vs fitted plot in @fig-9.
 From this plot we can see that there is a mostly linear trend with only a few straggling points throughout.
@@
 Also, the equal variance condition is met since there are only a couple of points that have concerningly high residuals.
-In the normal Q-Q plot, there is some tailing at both ends, but nothing to cause extreme concern so the normality condition is met.
+In the normal Q-Q plot in @fig-10, there is some tailing at both ends, but nothing to cause extreme concern so the normality condition is met.
 Since all three conditions are met, this model is valid for the purpose of prediction.
@@
 ```{r, echo = FALSE}
+#| echo: false
+#| label: fig-9
+#| warning: false
+#| fig-cap: "Residual Plot of Damping Fraction Model"
+#| fig-align: "center"
+#| fig-width: 7
+#| fig-height: 4
+
 plot(dampingModel)
+```
+
+```{r, echo = FALSE}
+#| echo: false
+#| label: fig-10
+#| warning: false
+#| fig-cap: "QQ Plot of Damping Fraction Model"
+#| fig-align: "center"
+#| fig-width: 7
+#| fig-height: 4
 qqnorm(resid(dampingModel))
@@
 
-Lastly, we must look at the random effects of our model.
+Lastly, we must look at the random effects of our model in @fig-11.
 Like the stiffness model, as the intercept increases, the slope decreases.
 Although the trend is not as linear as the stiffness model, it still supports our conclusion that a mixed model should be used.
-Again, we would expect to see that each of the foams starts at a similar dampint fraction and as time increases, there would be deviations based on the other predictors.
+Again, we would expect to see that each of the foams starts at a similar damping fraction and as time increases, there would be deviations based on the other predictors.
 This plot confirms this behavior.
@@
 ```{r, echo = FALSE}
+#| echo: false
+#| label: fig-11
+#| warning: false
+#| fig-cap: "Intercept Rand Effects vs. Slope Rand Effects"
+#| fig-align: "center"
+#| fig-width: 7
+#| fig-height: 4
+
 re <- ranef(dampingModel)$ID
@@
 
-*Succinct response to each question laid out in 1.2.  This a much shorter version of section 3, and focuses on conclusions rather than the analyses.*
-
-**Question 1:** We see that porosity, chemical index, and time have the most significant effects on damping and stiffness. Frequency also has a significant effect but only on damping.
+**Question 1:** We see that porosity, chemical index, and cycle number have the most significant effects on damping and stiffness. Frequency also has a significant effect, but only on damping.
 
@@
 The individual predictors will impact damping in a similar way but foam with a chemical index of 79 will see the greatest decay in damping.
-Overall, there appear to be significant impacts on the stiffness and damping values depending on chemical index and porosity
+Overall, there appear to be significant impacts on the stiffness and damping values depending on chemical index and porosity.
 
@@
 Our model for stiffness fraction, for example, consists of six possible trendlines, based on a foam’s porosity and chemical index, and most combinations of porosity and chemical index had fewer than ten foam samples in this experiment.
-We can confidently say that the coefficient estimates in both our models are non-zero, but the estimates could still be off.
+We can confidently say that the coefficient estimates in both our models are non-zero, but the estimates could still be off from their true values.
 We would need more observations to get a more reliable estimate of the true coefficients.
+
 Compounding this problem, other treatment levels like amplitude and initial stress were not used evenly in the experiment.
@@
 # Technical Appendix
-
-*Detailed information and a copy of code and or software results.  Additional graphs and supporting figures may also be placed in the appendix.*
 

### 2024-04-27 CS2 Final Draft.qmd
Add files via upload

--- 2024-04-25 draft1.Rmd
@@
----
-title: "Untitled"
-output: html_document
-date: "2024-04-24"
----
-
 ---
@@
 foam_data_scaled <- foam_data |>
-  filter(cycle <= 10000 & cycle > 0)
+  filter(cycle <= 10000 & cycle > 0) |>
+  mutate(dampingFrac = replace(dampingFrac, dampingFrac < 0, NA))
 
@@
 
-This project comes to us from Moria Foster, from the Lawrence Livermore National Laboratory.
+This project comes to us from Moira Foster, from the Lawrence Livermore National Laboratory.
 Foster has been running experiments on the Colorado School of Mines campus to study the fatigue responses of stiffness and damping in different varieties of foam.
@@
 
-Concerns regarding the data's adequacy for drawing cause-and-effect conclusions prompted us to reformat the data from wide to long, helping our statistical analysis by ensuring each observation represents one measurement.
-Additionally, to maintain consistency across observations, samples taken beyond 10,000 cycles were removed.
-This step ensured that our analysis focuses on the relevant time frame for evaluating foam responses.
-By doing this we are hoping that our EDA gives us helpful insights into how we can optimize foam padding in these military helmets.
+We reformatted the data from wide to long to ensure each observation represents one measurement, not one foam sample.
+Measurements taken beyond 10,000 cycles were removed to maintain consistency across observations, and because Moira identified this as the most relevant time frame.
 
-There were a few concerning observations with negative values of stiffness and/or damping.
-At Foster’s instruction, we converted the negative stiffness values to positive, and left the negative damping values how they were.
+There were a few concerning observations with negative values of stiffness, likely due to a fault in the measuring machine.
+At Foster’s instruction, we converted these negative stiffness values to positive.
+The transformed response dampingFrac also had five concerning observations, that were not just negative, but drastic outliers.
+These were changed to NA per Dr.
+Philtron's recommendation.
 
@@
 
-```{r}
-foam_Samples <- foam_data[ ,c("ID", "chemIndex", "porosity", "freq", "amp", "initialStrain", "initialStress")]
-foam_Samples <- unique(foam_Samples)
+| Porosity | **ChemIndex 79** | **ChemIndex 100** | **ChemIndex 121** |
+|---------:|:-----------------|:------------------|:------------------|
+| **81**   | 9                | 7                 | 6                 |
+| **71**   | 11               | 5                 | 8                 |
 
-count(foam_Samples, chemIndex)
-count(foam_Samples, porosity)
-count(foam_Samples, chemIndex, porosity)
-```
-
-The counts above show that unequal amounts of each foam type were tested.
-If a specific type of foam (like low-porosity foam with chemical index 2) has very few observations, it will be more difficult to make solid conclusions about this type of foam.
-Even the foam types with more samples still have relatively few, so making solid conclusions for these could be difficult as well.
+These counts show that a relatively small number of each type of foam were tested. The amounts are unequal as well, ranging from 5 to 11.
 
@@
 #| fig-align: "center"
-#| fig-width: 10
+#| fig-width: 4
 #| fig-height: 4
-#| 
+
+foam_Samples <- foam_data[ ,c("ID", "chemIndex", "porosity", "freq", "amp", "initialStrain", "initialStress")]
+foam_Samples <- unique(foam_Samples)
+
 ggplot(foam_Samples, aes(x=freq)) +
-  geom_histogram()
+  geom_histogram(bins=10)
 ```
@@
 #| warning: false
-#| fig-cap: "siffnessFrac vs cycle"
+#| fig-cap: "StiffnessFrac and DampingFrac vs Cycle"
 #| fig-align: "center"
@@
   geom_point() +
-  scale_y_continuous(limits = c(0.825, 1)) + 
   theme(legend.position="none")
 
-p3 <- ggplot(foam_data_scaled, aes(y = dampingFrac, x = cycle, col=factor(ID))) + 
-  geom_point() +
-  scale_y_continuous(limits = c(-20, 1)) +
-  theme(legend.position="none")
-
-combined <- p1 + p2 + p3
+combined <- p1 + p2
 
 combined
-
 ```
@@
 The dot colors in @fig-2 represent which foam sample they were measured from.
-In the first plot, stiffnessFrac seems to spread out in an umbrella shape, showing that as cycle increases the sample values not only change, but get further apart.
-So cycle number has a relationship with stiffnessFrac, but that relationship looks different for each foam sample.
+As cycle number increases, the stiffnessFrac and dampingFrac values not only change, but spread further apart.
+So stiffnessFrac and dampingFrac both appear to have a relationship with cycle, but that relationship looks different for each foam sample.
 
@@
 So cycle number looks like it has a negative relationship with dampingFrac, but that relationship still looks different for each foam sample.
-
-The second plot was scaled so that the scale was from 0.75-1, given that all but 13 dampingFrac observations were in this range.
-The observations outside of this range (included in the unscaled third plot) all occurred in the interval code N, which is the last interval code we have in our filtered data, most likely meaning that these foams broke in some way or the experiment failed.
 
@@
 
-In @fig-3, we logged the responses, and plotted them against the log of cycle. The relationships between log(cycle) and the logged responses look more linear than the relationships in @fig-2.
+In @fig-3, we logged the responses, and plotted them against the log of cycle.
+The relationships between log(cycle) and the logged responses look more linear than the relationships in @fig-2, so a linear model using these logged variables would fit better.
 
-In order to understand how log(stiffnessFrac) and log(dampingFrac) behave over log(cycle) given different frequencies, we also colored the samples in @fig-3 by frequency.
+In order to understand how log(stiffnessFrac) and log(dampingFrac) behave given different frequencies, we also colored the samples in @fig-3 by frequency.
 For log(stiffnessFrac), we see that higher frequencies have a more negative trend than lower frequencies.
@@
   geom_hline(yintercept = 0, linetype="dashed") +
-  ylim(-0.325, 0.15)
+  ylim(-0.325, 0.15) +
+  labs(title="ChemIndex 79, Porisity 81") +
+  theme(plot.title = element_text(size=25))
 sfc2 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 2 & foam_data$porosity == "high",], aes(x=log(cycle), y=log(stiffnessFrac))) +
@@
   geom_hline(yintercept = 0, linetype="dashed") +
-  ylim(-0.325, 0.15)
+  ylim(-0.325, 0.15) +
+  labs(title="ChemIndex 100, Porisity 81") +
+  theme(plot.title = element_text(size=25))
 sfc3 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 3 & foam_data$porosity == "high",], aes(x=log(cycle), y=log(stiffnessFrac))) +
@@
   geom_hline(yintercept = 0, linetype="dashed") +
-  ylim(-0.325, 0.15)
+  ylim(-0.325, 0.15) +
+  labs(title="ChemIndex 121, Porisity 81") +
+  theme(plot.title = element_text(size=25))
 sfc4 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 1 & foam_data$porosity == "low",], aes(x=log(cycle), y=log(stiffnessFrac))) +
@@
   geom_hline(yintercept = 0, linetype="dashed") +
-  ylim(-0.325, 0.15)
+  ylim(-0.325, 0.15) +
+  labs(title="ChemIndex 79, Porisity 71") +
+  theme(plot.title = element_text(size=25))
 sfc5 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 2 & foam_data$porosity == "low",], aes(x=log(cycle), y=log(stiffnessFrac))) +
@@
   geom_hline(yintercept = 0, linetype="dashed") +
-  ylim(-0.325, 0.15)
+  ylim(-0.325, 0.15) +
+  labs(title="ChemIndex 100, Porisity 71") +
+  theme(plot.title = element_text(size=25))
 sfc6 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 3 & foam_data$porosity == "low",], aes(x=log(cycle), y=log(stiffnessFrac))) +
@@
   geom_hline(yintercept = 0, linetype="dashed") +
-  ylim(-0.325, 0.15)
+  ylim(-0.325, 0.15) +
+  labs(title="ChemIndex 121, Porisity 71") +
+  theme(plot.title = element_text(size=25))
 
@@
 
-The plots in @fig-4 separate the log(stiffnessFrac) vs log(cycle) graph from above to show how the interaction between chemical index and porosity impact log(stiffnessFrac).
-The top row shows only high porosity and the bottom row shows only low porosity samples.
-From left to right, the columns show chemical index 79, 100 and 121.
+The plots in @fig-4 separate the log(stiffnessFrac) vs log(cycle) graph in @fig-3 to show how the interaction between chemical index and porosity impact log(stiffnessFrac).
 We can see that the chemical index 79 has many samples where log(stiffnessFrac) trends upwards, whereas chemical indices 100 and 121 mostly trend downwards.
@@
   geom_hline(yintercept = 0, linetype="dashed") +
-  ylim(-0.175, 0.125)
+  ylim(-0.175, 0.125) +
+  labs(title="ChemIndex 79, Porisity 81") +
+  theme(plot.title = element_text(size=25))
 dfc2 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 2 & foam_data_scaled$porosity == "high",], aes(x=log(cycle), y=log(dampingFrac))) +
@@
   geom_hline(yintercept = 0, linetype="dashed") +
-  ylim(-0.175, 0.125)
+  ylim(-0.175, 0.125) +
+  labs(title="ChemIndex 100, Porisity 81") +
+  theme(plot.title = element_text(size=25))
 dfc3 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 3 & foam_data_scaled$porosity == "high",], aes(x=log(cycle), y=log(dampingFrac))) +
@@
   geom_hline(yintercept = 0, linetype="dashed") +
-  ylim(-0.175, 0.125)
+  ylim(-0.175, 0.125) +
+  labs(title="ChemIndex 121, Porisity 81") +
+  theme(plot.title = element_text(size=25))
 dfc4 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 1 & foam_data_scaled$porosity == "low",], aes(x=log(cycle), y=log(dampingFrac))) +
@@
   geom_hline(yintercept = 0, linetype="dashed") +
-  ylim(-0.175, 0.125)
+  ylim(-0.175, 0.125) +
+  labs(title="ChemIndex 79, Porisity 71") +
+  theme(plot.title = element_text(size=25))
 dfc5 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 2 & foam_data_scaled$porosity == "low",], aes(x=log(cycle), y=log(dampingFrac))) +
@@
   geom_hline(yintercept = 0, linetype="dashed") +
-  ylim(-0.175, 0.125)
+  ylim(-0.175, 0.125) +
+  labs(title="ChemIndex 100, Porisity 71") +
+  theme(plot.title = element_text(size=25))
 dfc6 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 3 & foam_data_scaled$porosity == "low",], aes(x=log(cycle), y=log(dampingFrac))) +
@@
   geom_hline(yintercept = 0, linetype="dashed") +
-  ylim(-0.175, 0.125)
+  ylim(-0.175, 0.125) +
+  labs(title="ChemIndex 121, Porisity 71") +
+  theme(plot.title = element_text(size=25))
 
@@
 The plots in @fig-5 separate the log(dampingFrac) vs log(cycle) graph from above to show how the interaction between chemical index and porosity impact log(dampingFrac).
-The data shown in the rows and columns are the same as in @fig-4.
 We can see differences in the log(dampingFrac) trends between different chemical indices, so there is likely a relationship present, but this relationship looks less pronounced than it was for the log(stiffnessFrac) trends.
@@
 
+This is especially true for foams with especially few samples, like low-porosity foam with chemical index 100.
+Our model estimates regarding foams like this are even more liable to be wrong.
+So our models are not equally strong with all 6 types of foam.
+
 Compounding this problem, other treatment levels like amplitude and initial stress were not used evenly in the experiment.
@@
 With stiffnessFrac this was at bic = -640 and with dampingFrac this was at -520.
+
+```{r, include=TRUE}
+foam_Samples <- foam_data[ ,c("ID", "chemIndex", "porosity", "freq", "amp", "initialStrain", "initialStress")]
+foam_Samples <- unique(foam_Samples)
+
+count(foam_Samples, chemIndex)
+count(foam_Samples, porosity)
+count(foam_Samples, chemIndex, porosity)
+```

### 2024-04-29 CS2 Final Draft.qmd
Update CS2 Final Draft.qmd
updated #5

@@
 
-For further and future studies, we recommend that more foam samples are tested, so that there are more observations to work with.
-The same number of each type of foam should be used, and the treatments such as amplitude and initial stress should be assigned evenly, so each treatment level is fully tested on each type of foam.
-This will allow for more trustworthy models.
+For future studies and more reliable models, we recommend testing more foam samples to increase the number of observations that can be used.
+The same number of each type of foam should be used, and treatments such as amplitude and initial stress should be assigned evenly.
+This ensures each treatment level is fully tested on each type of foam.
+We would also recommend regulating the temperature of the testing environment to help provide the best testing conditions.
 

### 2024-04-29 CS2 Final Draft.qmd
Update CS2 Final Draft.qmd
updated #1.2

@@
 
-Explanatory Variables:
+**Explanatory Variables:**
 
@@
 
-Aside from cycle, all of these explanatory variables are constant across all measurements of an individual foam piece.
+*Caption: Table 1. Explanatory Variables*
 
-Response Variables:
+**Response Variables:**
 
@@
 |    stiffnessFrac     |  stiffness / first stiffness measurement|
-|     dampingFrac      |  tan(damping) / tan(first damping measurement)
+|     dampingFrac      |  tan(damping) / tan(first damping measurement)|
+
+*Caption: Table 2. Response Variables*
 

### 2024-04-29 CS2 Final Draft.qmd
Update CS2 Final Draft.qmd
Update CS2 Final Draft.qmd
updated #1.2 again

@@
 
-**Explanatory Variables:**
+\begin{center}
+\textbf{Explanatory Variables:}
+\end{center}
 
-|    Variable Name    |                   Description                                 |
-|:--------------------|:---------------------------------------------------------------------------------------------|
-|        ID            |  the unique identifier of the different foam pieces|
-|     chemIndex        |  the chemical makeup type of the foam. 1 = 79, 2 = 100, 3 = 121|
-|      porosity        |  the porosity level of the foam. Low = 71%, high = 81%|
-|       freq           |  the number of compressions per second (hz)|
-|        amp           |  the amplitude of the compressions|
-|    initialStrain     |  the initial measure of the strain of foam. (Originally labeled “strain”)|
-|    initialStress     |  the initial measure of the stress of the foam. (Originally labeled “stresskpa”)|
-|       cycle          |  the cycle number the measurement was taken at|
+\begin{center}
+\begin{tabular}{|l|l|}
+\hline
+\text{Variable Name} & \text{Description} \\
+\hline
+ID & the unique identifier of the different foam pieces \\
+chemIndex & the chemical makeup type of the foam. 1 = 79, 2 = 100, 3 = 121 \\
+porosity & the porosity level of the foam. Low = 71\%, high = 81\% \\
+freq & the number of compressions per second (hz) \\
+amp & the amplitude of the compressions \\
+initialStrain & the initial measure of the strain of foam. (Originally labeled ``strain'') \\
+initialStress & the initial measure of the stress of the foam. (Originally labeled ``stresskpa'') \\
+cycle & the cycle number the measurement was taken at \\
+\hline
+\end{tabular}
+\end{center}
 
-*Caption: Table 1. Explanatory Variables*
+\begin{center}
+\text{Table 1: Explanatory Variables}
+\end{center}
 
-**Response Variables:**
+\begin{center}
+\textbf{Response Variables}
+\end{center}
 
-|    Variable Name    |                   Description                                       |
-|:--------------------|:---------------------------------------------------------------------------------------------|
-|     stiffness        |  measurement of the foam’s stiffness. (Originally labeled “E1_MPA”)|
-|      damping         |  measurement of the foam’s damping. (Originally labeled “delta_rad”)|
-|    stiffnessFrac     |  stiffness / first stiffness measurement|
-|     dampingFrac      |  tan(damping) / tan(first damping measurement)|
+\begin{center}
+\begin{tabular}{|l|l|}
+\hline
+\text{Variable Name} & \text{Description} \\
+\hline
+stiffness & measurement of the foam’s stiffness. (Originally labeled ``E1\_MPA'') \\
+damping & measurement of the foam’s damping. (Originally labeled ``delta\_rad'') \\
+stiffnessFrac & stiffness / first stiffness measurement \\
+dampingFrac & tan(damping) / tan(first damping measurement) \\
+\hline
+\end{tabular}
+\end{center}
 
-*Caption: Table 2. Response Variables*
+\begin{center}
+\text{Table 2: Response Variables}
+\end{center}
 
@@
 
-$\log(\text{stiffnessFrac}) = \alpha_0 + \alpha_1*I(\text{porosity} = \text{low})) + \alpha_2*I(\text{chemIndex} = 2)$\newline
-$+\alpha_3*I(\text{chemIndex} = 3)+\alpha_4*I(\text{chemIndex} = 2)*I(\text{porosity} = low)$\newline
-$+\alpha_5*I(\text{chemIndex} = 3)*I(\text{porosity} = low)+\beta_0*\log(\text{cycle}) +  \beta_1*\log(\text{cycle})*I(\text{porosity} = \text{low})$\newline
-$+\beta_2*\log(\text{cycle})*I(\text{chemIndex} = 2)+\beta_3*\log(\text{cycle})*I(\text{chemIndex} = 2)*I(\text{porosity} = \text{low})$\newline
-$+\beta_4*\log(\text{cycle})*I(\text{chemIndex} = 3)*I(\text{porosity} = \text{low})+ u_i + v_i*\log(\text{cycle}) + \epsilon_{ij}$
+Level 1: $\log(\text{stiffnessFrac}) = \alpha_i + \beta_i*log(cycle)$
+Level 2: $\alpha_i = \alpha_0 + \alpha_1*\text{porosity} + \alpha_2*\text{chemIndex} + \alpha_3*\text{freq} + \alpha_4*\text{amp}$
+         $\beta_i = \beta_0 + \beta_1*\text{porosity} + \beta_2*\text{chemIndex} +\beta_3*\text{freq} + \beta_4*\text{amp}$
 
@@
 
-$\log(\text{dampingFrac}) = \alpha_0 + \alpha_1*I(\text{porosity} = \text{low}) + \alpha_2*I(\text{chemIndex} = 2) + \alpha_3*I(\text{chemIndex} = 3)$\newline
-$+\alpha_4*\text{freq} + \beta_0*\log(\text{cycle}) + \beta_1*(I(\text{porosity} = \text{low})*\log(\text{cycle}))$\newline
-$+\beta_3*(I(\text{chemIndex} = 2)*\log(\text{cycle})) + \beta_4*(I(\text{chemIndex} = 3)*\log(\text{cycle}))$\newline
-$+\beta_5*(\text{freq}*\log(\text{cycle})) + u_i + v_i*\log(\text{cycle}) + \epsilon_{ij}$.
+Level 1: $\log(\text{dampingFrac}) = \alpha_i + \beta_i*log(cycle)$
+Level 2: $\alpha_i = \alpha_0 + \alpha_1*\text{porosity} + \alpha_2*\text{chemIndex} + \alpha_3*\text{freq}$
+         $\beta_i = \beta_0 + \beta_1*\text{porosity} + \beta_2*\text{chemIndex} +\beta_3*\text{freq}$
 

### 2024-04-29 CS2 Final Draft.qmd
Update CS2 Final Draft.qmd

@@
 \hline
-\text{Variable Name} & \text{Description} \\
+\textbf{Variable Name} & \textbf{Description} \\
 \hline
-ID & the unique identifier of the different foam pieces \\
-chemIndex & the chemical makeup type of the foam. 1 = 79, 2 = 100, 3 = 121 \\
-porosity & the porosity level of the foam. Low = 71\%, high = 81\% \\
-freq & the number of compressions per second (hz) \\
-amp & the amplitude of the compressions \\
-initialStrain & the initial measure of the strain of foam. (Originally labeled ``strain'') \\
-initialStress & the initial measure of the stress of the foam. (Originally labeled ``stresskpa'') \\
-cycle & the cycle number the measurement was taken at \\
+ID & The unique identifier of the different foam pieces \\
+chemIndex & The chemical makeup type of the foam. 1 = 79, 2 = 100, 3 = 121 \\
+porosity & The porosity level of the foam. Low = 71\%, high = 81\% \\
+freq & The number of compressions per second (hz) \\
+amp & The amplitude of the compressions \\
+initialStrain & The initial measure of the strain of foam. (Originally labeled ``strain'') \\
+initialStress & The initial measure of the stress of the foam. (Originally labeled ``stresskpa'') \\
+cycle & The cycle number the measurement was taken at \\
 \hline
@@
 \hline
-\text{Variable Name} & \text{Description} \\
+\textbf{Variable Name} & \textbf{Description} \\
 \hline
-stiffness & measurement of the foam’s stiffness. (Originally labeled ``E1\_MPA'') \\
-damping & measurement of the foam’s damping. (Originally labeled ``delta\_rad'') \\
-stiffnessFrac & stiffness / first stiffness measurement \\
+stiffness & Measurement of the foam’s stiffness. (Originally labeled ``E1\_MPA'') \\
+damping & Measurement of the foam’s damping. (Originally labeled ``delta\_rad'') \\
+stiffnessFrac & Stiffness / first stiffness measurement \\
 dampingFrac & tan(damping) / tan(first damping measurement) \\

### 2024-04-29 CS2 Final Draft.qmd
Update CS2 Final Draft.qmd
Tweaks

@@
 At Foster’s instruction, we converted these negative stiffness values to positive.
-The transformed response dampingFrac also had five concerning observations, that were not just negative, but drastic outliers.
-These were changed to NA per Dr.
+Then we made the transformed response variable stiffnessFrac, which reports each stiffness measurement divided by that foam sample's initial stiffness measurement.
+
+There were many observations will negative values of damping, but we were instructed to leave these as they were.
+Then we made the transformed response variable dampingFrac, which reports the tangent of each damping measurement divided by the tangent of that foam sample's initial damping measurement.
+DampingFrac had five data points that were drastic outliers, but these were changed to NA per Dr.
 Philtron's recommendation.
@@
 
-Compounding this problem, other treatment levels like amplitude and initial stress were not used evenly in the experiment.
-For example, a few samples were compressed with abnormal amplitude.
-If this had an effect on those samples’ stiffness and damping, it could have a big impact on the coefficient estimates since there weren’t enough other samples to balance it out.
-It could be that amplitude or some other factor has a significant effect on stiffness and damping, but we could not assess that impact confidently because there were not enough observations at each treatment level.
-For this reason, a predictor that should have been included in the model could have been left out.
+Compounding this problem, other treatment levels were not used evenly in the experiment.
+As @fig-1 showed, fewer samples were tested at high frequencies.
+Our models' claims about the effects of high frequency should be taken with a grain of salt, because they are less reliable than its claims about low frequency.
+Additionally, if one variety of foam was never tested at high frequency, then the model's claims about that foam at high frequency are based on nothing, so should not be trusted at all.
+
+Our models could have also been thrown off by conditions we didn't consider.
+For example, we didn't use amplitude in our models, but it could be that high amplitude has an impact on foam decay, and we simply could not detect it because so few samples were tested at high amplitude.
+Then there are unmeasured experimental conditions, like room temperature and how many other people were using the laboratory.
+Since these conditions were not controlled for, they could have had an effect we were unable to detect.
 
@@
 
-For future studies and more reliable models, we recommend testing more foam samples to increase the number of observations that can be used.
-The same number of each type of foam should be used, and treatments such as amplitude and initial stress should be assigned evenly.
-This ensures each treatment level is fully tested on each type of foam.
-We would also recommend regulating the temperature of the testing environment to help provide the best testing conditions.
+For further studies, we recommend that more foam samples are tested, so that there are more observations to work with.
+The same number of each type of foam should be used, and the treatments such as amplitude and initial stress should be assigned evenly, so each treatment level is fully tested on each type of foam.
+This will allow for more trustworthy models.
 

### 2024-04-30 CS2 Final Draft.qmd
Update CS2 Final Draft.qmd

@@
 ```{r, echo = FALSE}
-summary(FactorModel)
+Anova(FactorModel)
 ```
@@
 
-```{r, echo = FALSE}
-#| echo: false
-#| label: fig-6
-#| warning: false
-#| fig-cap: "Residual Plot of Stiffness Fraction Model"
-#| fig-align: "center"
-#| fig-width: 7
-#| fig-height: 4
-
-plot(FactorModel)
-```
-
-```{r, echo = FALSE}
-#| echo: false
-#| label: fig-7
-#| warning: false
-#| fig-cap: "QQ Plot of Stiffness Fraction Model"
-#| fig-align: "center"
-#| fig-width: 7
-#| fig-height: 4
-
-qqnorm(resid(FactorModel))
-qqline(resid(FactorModel))
-```
-
 Lastly, given that we are using a mixed effects model, it is necessary to explore the random effects associated with the model.
@@
 #| echo: false
-#| label: fig-8
+#| label: fig-6
 #| warning: false
@@
 The summary of this model shows a high significance in almost all of our coefficients through small p-values.
-The only value that does not show a significant p-value is frequency, however we chose to keep that in the model because of how significant the interaction between frequency and log(cycle) is.
+The only value that does not show a significant p-value is chemIndex, however we chose to keep that in the model because of how significant the interaction between chemIndex and log(cycle) is.
 This model has multiple interactions: porosity and log(cycle), frequency and log(cycle), and chemical index and log(cycle).
 These interactions show that if all other variables are held constant, there is a combined effect of these interactions on the response variable.
-The REML is very negative, which shows that the model fits the data and there is minimal error.
 
 ```{r, echo = FALSE}
-summary(dampingModel)
+Anova(dampingModel)
 ```
 
-From the diagnostic plots, we first look at the residuals vs fitted plot in @fig-9.
-From this plot we can see that there is a mostly linear trend with only a few straggling points throughout.
-The linearity condition is met since there is no major trend to contradict it.
-Also, the equal variance condition is met since there are only a couple of points that have concerningly high residuals.
-In the normal Q-Q plot in @fig-10, there is some tailing at both ends, but nothing to cause extreme concern so the normality condition is met.
-Since all three conditions are met, this model is valid for the purpose of prediction.
+We used a residuals vs fitted plot and QQ plot to assess the conditions of our model. We found that this damping model meets all of these conditions which therefore we can analyze our model's results.
 
-```{r, echo = FALSE}
-#| echo: false
-#| label: fig-9
-#| warning: false
-#| fig-cap: "Residual Plot of Damping Fraction Model"
-#| fig-align: "center"
-#| fig-width: 7
-#| fig-height: 4
-
-plot(dampingModel)
-```
-
-```{r, echo = FALSE}
-#| echo: false
-#| label: fig-10
-#| warning: false
-#| fig-cap: "QQ Plot of Damping Fraction Model"
-#| fig-align: "center"
-#| fig-width: 7
-#| fig-height: 4
-qqnorm(resid(dampingModel))
-qqline(resid(dampingModel))
-```
-
-Lastly, we must look at the random effects of our model in @fig-11.
+Lastly, we must look at the random effects of our model in @fig-7.
 Like the stiffness model, as the intercept increases, the slope decreases.
@@
 #| echo: false
-#| label: fig-11
+#| label: fig-7
 #| warning: false
@@
 
+# EDA code
+
+# graph1
+foam_Samples <- foam_data[ ,c("ID", "chemIndex", "porosity", "freq", "amp", "initialStrain", "initialStress")]
+foam_Samples <- unique(foam_Samples)
+
+ggplot(foam_Samples, aes(x=freq)) +
+  geom_histogram(bins=10)
+
+# graph2
+p1 <- ggplot(foam_data_scaled, aes(y = stiffnessFrac, x = cycle, col=factor(ID))) +
+  geom_point() + 
+  theme(legend.position="none")
+
+p2 <- ggplot(foam_data_scaled, aes(y = dampingFrac, x = cycle, col=factor(ID))) + 
+  geom_point() +
+  theme(legend.position="none")
+
+combined <- p1 + p2
+
+combined
+
+# graph3
+sp1 <- ggplot(foam_data_scaled, aes(y = log(stiffnessFrac), x = log(cycle))) + geom_point(aes(color = freq))
+#sp2 <- ggplot(foam_data_scaled, aes(y = log(stiffnessFrac), x = log(cycle))) + geom_point(aes(color = factor(chemIndex)))
+#sp3 <- ggplot(foam_data_scaled, aes(y = log(stiffnessFrac), x = log(cycle))) + geom_point(aes(color = porosity))
+
+#combined <- sp1 + sp2 + sp3
+
+#combined
+
+dp1 <- ggplot(foam_data_scaled, aes(y = log(dampingFrac), x = log(cycle))) + geom_point(aes(color = freq))
+#dp2 <- ggplot(foam_data_scaled, aes(y = log(dampingFrac), x = log(cycle))) + geom_point(aes(color = factor(chemIndex)))
+#dp3 <- ggplot(foam_data_scaled, aes(y = log(dampingFrac), x = log(cycle))) + geom_point(aes(color = porosity))
+
+#combined <- dp1 + dp2 + dp3
+
+#combined
+
+combined <- sp1 + dp1
+combined
+
+# graph4
+sfc1 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 1 & foam_data$porosity == "high",], aes(x=log(cycle), y=log(stiffnessFrac))) +
+  geom_point(color="red", size=3) + 
+  geom_hline(yintercept = 0, linetype="dashed") +
+  ylim(-0.325, 0.15) +
+  labs(title="ChemIndex 79, Porisity 81") +
+  theme(plot.title = element_text(size=25))
+sfc2 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 2 & foam_data$porosity == "high",], aes(x=log(cycle), y=log(stiffnessFrac))) +
+  geom_point(color="green", size=3) + 
+  geom_hline(yintercept = 0, linetype="dashed") +
+  ylim(-0.325, 0.15) +
+  labs(title="ChemIndex 100, Porisity 81") +
+  theme(plot.title = element_text(size=25))
+sfc3 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 3 & foam_data$porosity == "high",], aes(x=log(cycle), y=log(stiffnessFrac))) +
+  geom_point(color="cyan2", size=3) + 
+  geom_hline(yintercept = 0, linetype="dashed") +
+  ylim(-0.325, 0.15) +
+  labs(title="ChemIndex 121, Porisity 81") +
+  theme(plot.title = element_text(size=25))
+sfc4 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 1 & foam_data$porosity == "low",], aes(x=log(cycle), y=log(stiffnessFrac))) +
+  geom_point(color="red3", size=3) + 
+  geom_hline(yintercept = 0, linetype="dashed") +
+  ylim(-0.325, 0.15) +
+  labs(title="ChemIndex 79, Porisity 71") +
+  theme(plot.title = element_text(size=25))
+sfc5 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 2 & foam_data$porosity == "low",], aes(x=log(cycle), y=log(stiffnessFrac))) +
+  geom_point(color="green4", size=3) + 
+  geom_hline(yintercept = 0, linetype="dashed") +
+  ylim(-0.325, 0.15) +
+  labs(title="ChemIndex 100, Porisity 71") +
+  theme(plot.title = element_text(size=25))
+sfc6 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 3 & foam_data$porosity == "low",], aes(x=log(cycle), y=log(stiffnessFrac))) +
+  geom_point(col="blue", size=3) + 
+  geom_hline(yintercept = 0, linetype="dashed") +
+  ylim(-0.325, 0.15) +
+  labs(title="ChemIndex 121, Porisity 71") +
+  theme(plot.title = element_text(size=25))
+
+combined <- sfc1 + sfc2 + sfc3 + sfc4 + sfc5 + sfc6 + plot_layout(ncol = 3)
+combined
+
+# graph5
+dfc1 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 1 & foam_data_scaled$porosity == "high",], aes(x=log(cycle), y=log(dampingFrac))) +
+  geom_point(color="red", size=3) + 
+  geom_hline(yintercept = 0, linetype="dashed") +
+  ylim(-0.175, 0.125) +
+  labs(title="ChemIndex 79, Porisity 81") +
+  theme(plot.title = element_text(size=25))
+dfc2 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 2 & foam_data_scaled$porosity == "high",], aes(x=log(cycle), y=log(dampingFrac))) +
+  geom_point(color="green", size=3) + 
+  geom_hline(yintercept = 0, linetype="dashed") +
+  ylim(-0.175, 0.125) +
+  labs(title="ChemIndex 100, Porisity 81") +
+  theme(plot.title = element_text(size=25))
+dfc3 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 3 & foam_data_scaled$porosity == "high",], aes(x=log(cycle), y=log(dampingFrac))) +
+  geom_point(color="cyan2", size=3) + 
+  geom_hline(yintercept = 0, linetype="dashed") +
+  ylim(-0.175, 0.125) +
+  labs(title="ChemIndex 121, Porisity 81") +
+  theme(plot.title = element_text(size=25))
+dfc4 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 1 & foam_data_scaled$porosity == "low",], aes(x=log(cycle), y=log(dampingFrac))) +
+  geom_point(color="red3", size=3) + 
+  geom_hline(yintercept = 0, linetype="dashed") +
+  ylim(-0.175, 0.125) +
+  labs(title="ChemIndex 79, Porisity 71") +
+  theme(plot.title = element_text(size=25))
+dfc5 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 2 & foam_data_scaled$porosity == "low",], aes(x=log(cycle), y=log(dampingFrac))) +
+  geom_point(color="green4", size=3) + 
+  geom_hline(yintercept = 0, linetype="dashed") +
+  ylim(-0.175, 0.125) +
+  labs(title="ChemIndex 100, Porisity 71") +
+  theme(plot.title = element_text(size=25))
+dfc6 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 3 & foam_data_scaled$porosity == "low",], aes(x=log(cycle), y=log(dampingFrac))) +
+  geom_point(col="blue", size=3) + 
+  geom_hline(yintercept = 0, linetype="dashed") +
+  ylim(-0.175, 0.125) +
+  labs(title="ChemIndex 121, Porisity 71") +
+  theme(plot.title = element_text(size=25))
+
+combined <- dfc1 + dfc2 + dfc3 + dfc4 + dfc5 + dfc6 + plot_layout(ncol = 3)
+combined
+
 # summary of stiffnessFrac model 
@@
 plot(re)
+```
+
+```{r, include=TRUE}
+foam_Samples <- foam_data[ ,c("ID", "chemIndex", "porosity", "freq", "amp", "initialStrain", "initialStress")]
+foam_Samples <- unique(foam_Samples)
+
+count(foam_Samples, chemIndex)
+count(foam_Samples, porosity)
+count(foam_Samples, chemIndex, porosity)
 ```
@@
 
-@fig-example1 and @fig-example2 show the plots that we used to pick our predictors for our model.
-Since our main concerns for this analysis were with chemIndex and porosity, we chose the predictors with the highest bic starting at where those two predictors became significant.
-With stiffnessFrac this was at bic = -640 and with dampingFrac this was at -520.
+```{r, echo = FALSE}
+#| echo: false
+#| label: fig-example3
+#| warning: false
+#| fig-cap: "Residual Plot of Stiffness Fraction Model"
+#| fig-align: "center"
+#| fig-width: 7
+#| fig-height: 4
 
-```{r, include=TRUE}
-foam_Samples <- foam_data[ ,c("ID", "chemIndex", "porosity", "freq", "amp", "initialStrain", "initialStress")]
-foam_Samples <- unique(foam_Samples)
+plot(FactorModel)
+```
 
-count(foam_Samples, chemIndex)
-count(foam_Samples, porosity)
-count(foam_Samples, chemIndex, porosity)
+```{r, echo = FALSE}
+#| echo: false
+#| label: fig-example4
+#| warning: false
+#| fig-cap: "QQ Plot of Stiffness Fraction Model"
+#| fig-align: "center"
+#| fig-width: 7
+#| fig-height: 4
+
+qqnorm(resid(FactorModel))
+qqline(resid(FactorModel))
 ```
+
+```{r, echo = FALSE}
+#| echo: false
+#| label: fig-example5
+#| warning: false
+#| fig-cap: "Residual Plot of Damping Fraction Model"
+#| fig-align: "center"
+#| fig-width: 7
+#| fig-height: 4
+
+plot(dampingModel)
+```
+
+```{r, echo = FALSE}
+#| echo: false
+#| label: fig-example6
+#| warning: false
+#| fig-cap: "QQ Plot of Damping Fraction Model"
+#| fig-align: "center"
+#| fig-width: 7
+#| fig-height: 4
+qqnorm(resid(dampingModel))
+qqline(resid(dampingModel))
+```

### 2024-04-30 CS2 Final Draft.qmd
Update CS2 Final Draft.qmd

@@
 |:---------------------------------|:------------------------------------|
-            porosity               |             porosity
            chemIndex               |             chemIndex
-           log(cycle)              |               freq
-       porosity:chemIndex          |             log(cycle)
-       chemIndex:log(cycle)        |          porosity:log(cycle)
-       porosity:log(cycle)         |         chemIndex:log(cycle)
-  porosity:chemIndex:log(cycle)    |            freq:log(cycle)
+             freq                  |               freq
+          log(cycle)               |            log(cycle)
+              amp                  |             porosity
+       chemIndex:log(cycle)        |         chemIndex:log(cycle)
+        freq:log(cycle)            |           freq:log(cycle)
+        amp:log(cycle)             |         porosity:log(cycle)
 
@@
 Level 1: $\log(\text{stiffnessFrac}) = \alpha_i + \beta_i*log(cycle)$
-Level 2: $\alpha_i = \alpha_0 + \alpha_1*\text{porosity} + \alpha_2*\text{chemIndex} + \alpha_3*\text{freq} + \alpha_4*\text{amp}$
-         $\beta_i = \beta_0 + \beta_1*\text{porosity} + \beta_2*\text{chemIndex} +\beta_3*\text{freq} + \beta_4*\text{amp}$
+Level 2: $\alpha_i = \alpha_0 + \alpha_1*\text{chemIndex} + \alpha_2*\text{freq} + \alpha_3*\text{amp}$
+         $\beta_i = \beta_0 + \beta_1*\text{chemIndex} +\beta_2*\text{freq} + \beta_3*\text{amp}$
 

### 2024-05-02 CS2 Final Draft.qmd
Update CS2 Final Draft.qmd

@@
 
-FactorModel <- lmer(log(stiffnessFrac) ~  porosity + factor(chemIndex) + factor(chemIndex):porosity +  log(cycle) +  factor(chemIndex):log(cycle) + porosity:log(cycle) + factor(chemIndex):porosity:log(cycle) + (1 + log(cycle)|ID), data = foam_data_scaled)
+FactorModel <- lmer(log(stiffnessFrac) ~  factor(chemIndex) + freq + amp +  log(cycle) +  factor(chemIndex):log(cycle) + freq:log(cycle) + amp:log(cycle) + (1 + log(cycle)|ID), data = foam_data_scaled)
 
@@
 
-The summary of this model shows a high significance in almost all of our coefficients through small p-values.
+The summary of this model shows a high significance in almost all of our coefficients through small p-values (less than 0.05).
+These small p-values suggest that these variables are significant predictors of our response variable.
 The only value that does not show a significant p-value is chemIndex, however we chose to keep that in the model because of how significant the interaction between chemIndex and log(cycle) is.
@@
 
-**Question 1:** We see that porosity, chemical index, and cycle number have the most significant effects on damping and stiffness. Frequency also has a significant effect, but only on damping.
+**Question 1:** We see that chemical index, frequency, and cycle number have the most significant effects on both damping and stiffness.
+However porosity has an effect on damping and amplitude as an effect on stiffness.
 

### 2024-05-06 CS2 Final Draft.qmd
Tweak CS2 Final Draft.qmd

@@
 author: 'Katie Greene, Ava Harman, Chris Seery, Ian Stonecypher'
-date: 'April, 2024'
+date: 'May, 2024'
 format: pdf
@@
 
-We reformatted the data from wide to long to ensure each observation represents one measurement, not one foam sample.
+We reformatted the data from wide to long to ensure each observation represented one measurement, not one foam sample.
 Measurements taken beyond 10,000 cycles were removed to maintain consistency across observations, and because Moira identified this as the most relevant time frame.
@@
 
-There were many observations will negative values of damping, but we were instructed to leave these as they were.
+There were many observations with negative values of damping, but we were instructed to leave these as they were.
 Then we made the transformed response variable dampingFrac, which reports the tangent of each damping measurement divided by the tangent of that foam sample's initial damping measurement.
@@
 
-The histogram in @fig-1 shows the distribution of compression frequencies among the foam samples.
-The uneven distribution suggests that certain compression frequencies were tested more frequently than others.
-This non-uniformity in sample distribution might impact the reliability of conclusions drawn from the analysis, especially if frequencies with fewer samples have significant effects on the response variables.
+The histogram in @fig-1 shows the distribution of compression frequencies among the foam samples revealing that certain compression frequencies were tested more frequently than others.
+These non-uniformities in sample distribution might impact the reliability of conclusions drawn from the analysis.
 
@@
 So stiffnessFrac and dampingFrac both appear to have a relationship with cycle, but that relationship looks different for each foam sample.
-
-In the second plot, there seems to be a general downward stretching of dampingFrac values as cycle increases, also showing us that as cycle increases sample values seem to spread out.
-So cycle number looks like it has a negative relationship with dampingFrac, but that relationship still looks different for each foam sample.
 
@@
 
-Level 1: $\log(\text{stiffnessFrac}) = \alpha_i + \beta_i*log(cycle)$
-Level 2: $\alpha_i = \alpha_0 + \alpha_1*\text{chemIndex} + \alpha_2*\text{freq} + \alpha_3*\text{amp}$
+**Level 1**: $\log(\text{stiffnessFrac}) = \alpha_i + \beta_i*log(cycle)$
+
+**Level 2**: $\alpha_i = \alpha_0 + \alpha_1*\text{chemIndex} + \alpha_2*\text{freq} + \alpha_3*\text{amp}$
+
 $\beta_i = \beta_0 + \beta_1*\text{chemIndex} +\beta_2*\text{freq} + \beta_3*\text{amp}$
@@
 We also assessed the model to see how it met the conditions of a predictive model.
-From @fig-6, we see that there appears to be some sinusoidal behavior at the beginning and the middle of the shape has a larger range than the rest of the plot.
+From @fig-example3 (see Appendix), we see that there appears to be some sinusoidal behavior at the beginning and the middle of the shape has a larger range than the rest of the plot.
 Since the range of the plot is so small, we can conclude that the equal variance condition is being met.
 We also do not see any flaring or nonlinear movement so the linearity condition is also being met.
-The QQ plot in #fig-7 shows some tailing on the ends but nothing too concerning.
+The QQ plot in @fig-example4 (see Appendix) shows some tailing on the ends but nothing too concerning.
 This concludes that all conditions are being met so this model is valid for the purpose of prediction.
@@
 Lastly, given that we are using a mixed effects model, it is necessary to explore the random effects associated with the model.
-From @fig-8, we notice that as the intercept increases, the slope decreases.
+From @fig-6, we notice that as the intercept increases, the slope decreases.
 This is a very defined trend and supports our conclusion that a mixed model should be used.
@@
 
-Level 1: $\log(\text{dampingFrac}) = \alpha_i + \beta_i*log(cycle)$
-Level 2: $\alpha_i = \alpha_0 + \alpha_1*\text{porosity} + \alpha_2*\text{chemIndex} + \alpha_3*\text{freq}$
+**Level 1**: $\log(\text{dampingFrac}) = \alpha_i + \beta_i*log(cycle)$
+
+**Level 2**: $\alpha_i = \alpha_0 + \alpha_1*\text{porosity} + \alpha_2*\text{chemIndex} + \alpha_3*\text{freq}$
+
 $\beta_i = \beta_0 + \beta_1*\text{porosity} + \beta_2*\text{chemIndex} +\beta_3*\text{freq}$
@@
 These small p-values suggest that these variables are significant predictors of our response variable.
-The only value that does not show a significant p-value is chemIndex, however we chose to keep that in the model because of how significant the interaction between chemIndex and log(cycle) is.
+The only value that does not show a significant p-value is chemIndex, which we chose to keep because the interaction between chemIndex and log(cycle) is significant.
 This model has multiple interactions: porosity and log(cycle), frequency and log(cycle), and chemical index and log(cycle).
@@
 
-We used a residuals vs fitted plot and QQ plot to assess the conditions of our model. We found that this damping model meets all of these conditions which therefore we can analyze our model's results.
+We used a residuals vs fitted plot and QQ plot to assess the conditions of our model (see @fig-example5 and @fig-example6, in the Appendix).
+We found that this damping model meets all of these conditions which therefore we can analyze our model's results.
 
@@
 One concern we have about our conclusions is that they were made based on relatively few observations.
-Our model for stiffness fraction, for example, consists of six possible trendlines, based on a foam’s porosity and chemical index, and most combinations of porosity and chemical index had fewer than ten foam samples in this experiment.
-We can confidently say that the coefficient estimates in both our models are non-zero, but the estimates could still be off from their true values.
+Our model for stiffness fraction, for example, depends on chemical index, frequency, and amplitude, and only 12 samples with chemical index 100 were tested.
+There are too many possible combinations of frequency and amplitude to be explored thoroughly with just 12 samples.
+In cases like this, our models' coefficient estimates are based on very little information.
+The p-values gave us confidence that the coefficients were not zero, but our estimates could still be far off from their true values.
 We would need more observations to get a more reliable estimate of the true coefficients.
 
-This is especially true for foams with especially few samples, like low-porosity foam with chemical index 100.
-Our model estimates regarding foams like this are even more liable to be wrong.
-So our models are not equally strong with all 6 types of foam.
+A similar concern comes from foam types and testing conditions that were used less often, such as high frequency.
+The reliability of our models is already limited, and for these cases they are even less reliable.
+Additionally, if one variety of foam was never tested at high frequency, then the model's claims about that foam at high frequency are based on nothing, and should not be trusted at all.
+So not all information this model gives on what effects stiffness and damping in foam is equally trustworthy.
 
-Compounding this problem, other treatment levels were not used evenly in the experiment.
-As @fig-1 showed, fewer samples were tested at high frequencies.
-Our models' claims about the effects of high frequency should be taken with a grain of salt, because they are less reliable than its claims about low frequency.
-Additionally, if one variety of foam was never tested at high frequency, then the model's claims about that foam at high frequency are based on nothing, so should not be trusted at all.
-
-Our models could have also been thrown off by conditions we didn't consider.
-For example, we didn't use amplitude in our models, but it could be that high amplitude has an impact on foam decay, and we simply could not detect it because so few samples were tested at high amplitude.
-Then there are unmeasured experimental conditions, like room temperature and how many other people were using the laboratory.
+Our models could have also been thrown off by conditions we didn't consider, such as room temperature and how many other people were using the laboratory.
 Since these conditions were not controlled for, they could have had an effect we were unable to detect.
@@
 For further studies, we recommend that more foam samples are tested, so that there are more observations to work with.
-The same number of each type of foam should be used, and the treatments such as amplitude and initial stress should be assigned evenly, so each treatment level is fully tested on each type of foam.
+The same number of each type of foam should be used, and the treatments such as amplitude and initial stress should be assigned evenly and randomly, so each treatment level is fully tested on each type of foam.
 This will allow for more trustworthy models.

### 2024-05-06 CS2 Final Draft.qmd
Edit Appendix

@@
 
+\begin{center}
+\text{Table 3: Foam Sample Type Counts}
+\end{center}
+
 These counts show that a relatively small number of each type of foam were tested. The amounts are unequal as well, ranging from 5 to 11.
@@
 As cycle number increases, the stiffnessFrac and dampingFrac values not only change, but spread further apart.
-So stiffnessFrac and dampingFrac both appear to have a relationship with cycle, but that relationship looks different for each foam sample.
+StiffnessFrac and dampingFrac both appear to have a relationship with cycle, but that relationship looks different for each foam sample.
 
@@
 #| warning: false
-#| fig-cap: "log(stiffnessFrac) vs log(cycle) by freq, and log(dampingFrac) vs log(cycle) by freq"
+#| fig-cap: "Log(stiffnessFrac) vs Log(cycle) and Log(dampingFrac) vs Log(cycle), Colored by Freq"
 #| fig-align: "center"
@@
 For log(dampingFrac), we see that lower frequencies have a more negative trend than higher frequencies.
-So frequency looks related to both stiffnessFrac and dampingFrac.
+Frequency looks related to both stiffnessFrac and dampingFrac.
 
@@
 #| warning: false
-#| fig-cap: "log(stiffnessFrac) vs log(cycle) by chemIndex and porosity"
+#| fig-cap: "Log(stiffnessFrac) vs Log(cycle), Separated by ChemIndex and Porosity"
 #| fig-align: "center"
@@
 #| warning: false
-#| fig-cap: "log(dampingFrac) vs log(cycle) for chemIndex and porosity"
+#| fig-cap: "Log(dampingFrac) vs Log(cycle), Separated by ChemIndex and Porosity"
 #| fig-align: "center"
@@
         amp:log(cycle)             |         porosity:log(cycle)
+
+\begin{center}
+\text{Table 4: Predictors Selected For Each Model}
+\end{center}
 
@@
 
-### R Script
-```{r ref.label=c('front-matter', 'fig-example1','tbl-summarystats', 'fig-example2')}
+### Data Wrangling
+
+```{r}
 #| echo: true
 #| eval: false
+knitr::opts_chunk$set(echo = TRUE)
+
+library(patchwork)
+library("readxl")
+library(ggplot2)
+library(tidyverse)
+
+foamDataRaw <- read_excel("DMA_Fatigue_Data_from_Team_Wendy_Polyurethane
+                          _LLNL-MI-860456.xlsx")
+names(foamDataRaw) <- foamDataRaw[3,]
+foamDataRaw <- foamDataRaw[-c(1:3),] # remove rows without data
+foamDataRaw <- foamDataRaw[,-c(62:64)] # remove duplicate columns
+
+foamDataIntermediate <- foamDataRaw |>
+  separate(col=SampleCode, into=c("Chemical", "chemIndex", "porosity", "ID")) |>
+  select(!Chemical) |>
+  mutate_at(vars(chemIndex), as.character) |>
+  mutate_at(vars(porosity), as.character) |>
+  mutate(chemIndex = dplyr::recode(chemIndex, "79"="1", "100"="2", "121"="3")) |>
+  mutate(porosity = dplyr::recode(porosity, "71"="low", "81"="high")) |>
+  rename("initialStrain"="strain") |>
+  rename("initialStress"="stresskpa") |>
+  mutate_at(c("Cyc_A", "Cyc_B", "Cyc_C", "Cyc_D", "Cyc_E", "Cyc_F", "Cyc_G", "Cyc_H", "Cyc_I", 
+              "Cyc_J", "Cyc_K", "Cyc_L", "Cyc_M", "Cyc_N", "E1_MPA_Cyc_A", "E1_MPA_Cyc_B", 
+              "E1_MPA_Cyc_C", "E1_MPA_Cyc_D", "E1_MPA_Cyc_E", "E1_MPA_Cyc_F", "E1_MPA_Cyc_G", 
+              "E1_MPA_Cyc_H", "E1_MPA_Cyc_I", "E1_MPA_Cyc_J", "E1_MPA_Cyc_K", "E1_MPA_Cyc_L", 
+              "E1_MPA_Cyc_M", "E1_MPA_Cyc_N", "delta_rad_Cyc_A", "delta_rad_Cyc_B", 
+              "delta_rad_Cyc_C", "delta_rad_Cyc_D", "delta_rad_Cyc_E", "delta_rad_Cyc_F", 
+              "delta_rad_Cyc_G", "delta_rad_Cyc_H", "delta_rad_Cyc_I", "delta_rad_Cyc_J", 
+              "delta_rad_Cyc_K", "delta_rad_Cyc_L", "delta_rad_Cyc_M", "delta_rad_Cyc_N"), 
+            as.numeric) |>
+  mutate_at(c("E1_MPA_Cyc_A", "E1_MPA_Cyc_B", "E1_MPA_Cyc_C", "E1_MPA_Cyc_D", "E1_MPA_Cyc_E", 
+              "E1_MPA_Cyc_F", "E1_MPA_Cyc_G", "E1_MPA_Cyc_H", "E1_MPA_Cyc_I", "E1_MPA_Cyc_J", 
+              "E1_MPA_Cyc_K", "E1_MPA_Cyc_L", "E1_MPA_Cyc_M", "E1_MPA_Cyc_N"), abs)
+
+#These lines add columns for Rate Of Change of stiffness and damping.
+foamDataIntermediate <- foamDataIntermediate |>
+  mutate(stiffness_ROC_A = NA) |>
+  mutate(stiffness_ROC_B = (E1_MPA_Cyc_B - E1_MPA_Cyc_A)/(Cyc_B - Cyc_A))|>
+  mutate(stiffness_ROC_C = (E1_MPA_Cyc_C - E1_MPA_Cyc_B)/(Cyc_C - Cyc_B))|>
+  mutate(stiffness_ROC_D = (E1_MPA_Cyc_D - E1_MPA_Cyc_C)/(Cyc_D - Cyc_C))|>
+  mutate(stiffness_ROC_E = (E1_MPA_Cyc_E - E1_MPA_Cyc_D)/(Cyc_E - Cyc_D))|>
+  mutate(stiffness_ROC_F = (E1_MPA_Cyc_F - E1_MPA_Cyc_E)/(Cyc_F - Cyc_E))|>
+  mutate(stiffness_ROC_G = (E1_MPA_Cyc_G - E1_MPA_Cyc_F)/(Cyc_G - Cyc_F))|>
+  mutate(stiffness_ROC_H = (E1_MPA_Cyc_H - E1_MPA_Cyc_G)/(Cyc_H - Cyc_G))|>
+  mutate(stiffness_ROC_I = (E1_MPA_Cyc_I - E1_MPA_Cyc_H)/(Cyc_I - Cyc_H))|>
+  mutate(stiffness_ROC_J = (E1_MPA_Cyc_J - E1_MPA_Cyc_I)/(Cyc_J - Cyc_I))|>
+  mutate(stiffness_ROC_K = (E1_MPA_Cyc_K - E1_MPA_Cyc_J)/(Cyc_K - Cyc_J))|>
+  mutate(stiffness_ROC_L = (E1_MPA_Cyc_L - E1_MPA_Cyc_K)/(Cyc_L - Cyc_K))|>
+  mutate(stiffness_ROC_M = (E1_MPA_Cyc_M - E1_MPA_Cyc_L)/(Cyc_M - Cyc_L))|>
+  mutate(stiffness_ROC_N = (E1_MPA_Cyc_N - E1_MPA_Cyc_M)/(Cyc_N - Cyc_M))|>
+  mutate(stiffness_frac_A = E1_MPA_Cyc_A / E1_MPA_Cyc_A) |>
+  mutate(stiffness_frac_B = E1_MPA_Cyc_B / E1_MPA_Cyc_A) |>
+  mutate(stiffness_frac_C = E1_MPA_Cyc_C / E1_MPA_Cyc_A) |>
+  mutate(stiffness_frac_D = E1_MPA_Cyc_D / E1_MPA_Cyc_A) |>
+  mutate(stiffness_frac_E = E1_MPA_Cyc_E / E1_MPA_Cyc_A) |>
+  mutate(stiffness_frac_F = E1_MPA_Cyc_F / E1_MPA_Cyc_A) |>
+  mutate(stiffness_frac_G = E1_MPA_Cyc_G / E1_MPA_Cyc_A) |>
+  mutate(stiffness_frac_H = E1_MPA_Cyc_H / E1_MPA_Cyc_A) |>
+  mutate(stiffness_frac_I = E1_MPA_Cyc_I / E1_MPA_Cyc_A) |>
+  mutate(stiffness_frac_J = E1_MPA_Cyc_J / E1_MPA_Cyc_A) |>
+  mutate(stiffness_frac_K = E1_MPA_Cyc_K / E1_MPA_Cyc_A) |>
+  mutate(stiffness_frac_L = E1_MPA_Cyc_L / E1_MPA_Cyc_A) |>
+  mutate(stiffness_frac_M = E1_MPA_Cyc_M / E1_MPA_Cyc_A) |>
+  mutate(stiffness_frac_N = E1_MPA_Cyc_N / E1_MPA_Cyc_A) |>
+  mutate(damping_frac_A = tan(delta_rad_Cyc_A) / tan(delta_rad_Cyc_A)) |>
+  mutate(damping_frac_B = tan(delta_rad_Cyc_B) / tan(delta_rad_Cyc_A)) |>
+  mutate(damping_frac_C = tan(delta_rad_Cyc_C) / tan(delta_rad_Cyc_A)) |>
+  mutate(damping_frac_D = tan(delta_rad_Cyc_D) / tan(delta_rad_Cyc_A)) |>
+  mutate(damping_frac_E = tan(delta_rad_Cyc_E) / tan(delta_rad_Cyc_A)) |>
+  mutate(damping_frac_F = tan(delta_rad_Cyc_F) / tan(delta_rad_Cyc_A)) |>
+  mutate(damping_frac_G = tan(delta_rad_Cyc_G) / tan(delta_rad_Cyc_A)) |>
+  mutate(damping_frac_H = tan(delta_rad_Cyc_H) / tan(delta_rad_Cyc_A)) |>
+  mutate(damping_frac_I = tan(delta_rad_Cyc_I) / tan(delta_rad_Cyc_A)) |>
+  mutate(damping_frac_J = tan(delta_rad_Cyc_J) / tan(delta_rad_Cyc_A)) |>
+  mutate(damping_frac_K = tan(delta_rad_Cyc_K) / tan(delta_rad_Cyc_A)) |>
+  mutate(damping_frac_L = tan(delta_rad_Cyc_L) / tan(delta_rad_Cyc_A)) |>
+  mutate(damping_frac_M = tan(delta_rad_Cyc_M) / tan(delta_rad_Cyc_A)) |>
+  mutate(damping_frac_N = delta_rad_Cyc_N / delta_rad_Cyc_A) |>mutate(damping_ROC_A = NA) |>
+  mutate(damping_ROC_B = (delta_rad_Cyc_B - delta_rad_Cyc_A)/(Cyc_B - Cyc_A))|>
+  mutate(damping_ROC_C = (delta_rad_Cyc_C - delta_rad_Cyc_B)/(Cyc_C - Cyc_B))|>
+  mutate(damping_ROC_D = (delta_rad_Cyc_D - delta_rad_Cyc_C)/(Cyc_D - Cyc_C))|>
+  mutate(damping_ROC_E = (delta_rad_Cyc_E - delta_rad_Cyc_D)/(Cyc_E - Cyc_D))|>
+  mutate(damping_ROC_F = (delta_rad_Cyc_F - delta_rad_Cyc_E)/(Cyc_F - Cyc_E))|>
+  mutate(damping_ROC_G = (delta_rad_Cyc_G - delta_rad_Cyc_F)/(Cyc_G - Cyc_F))|>
+  mutate(damping_ROC_H = (delta_rad_Cyc_H - delta_rad_Cyc_G)/(Cyc_H - Cyc_G))|>
+  mutate(damping_ROC_I = (delta_rad_Cyc_I - delta_rad_Cyc_H)/(Cyc_I - Cyc_H))|>
+  mutate(damping_ROC_J = (delta_rad_Cyc_J - delta_rad_Cyc_I)/(Cyc_J - Cyc_I))|>
+  mutate(damping_ROC_K = (delta_rad_Cyc_K - delta_rad_Cyc_J)/(Cyc_K - Cyc_J))|>
+  mutate(damping_ROC_L = (delta_rad_Cyc_L - delta_rad_Cyc_K)/(Cyc_L - Cyc_K))|>
+  mutate(damping_ROC_M = (delta_rad_Cyc_M - delta_rad_Cyc_L)/(Cyc_M - Cyc_L))|>
+  mutate(damping_ROC_N = (delta_rad_Cyc_N - delta_rad_Cyc_M)/(Cyc_N - Cyc_M))
+
+# These lines lengthen the data by putting each measurement interval in its own data frame, 
+#and row-binding them together.
+foamDataA <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_A, E1_MPA_Cyc_A, 
+                    stiffness_ROC_A, stiffness_frac_A, delta_rad_Cyc_A, damping_ROC_A, 
+                    damping_frac_A, NVP_Cyc_A) |>
+  rename("cycle"="Cyc_A") |>
+  rename("stiffness"="E1_MPA_Cyc_A") |>
+  rename("damping"="delta_rad_Cyc_A") |>
+  rename("NVP"="NVP_Cyc_A") |>
+  mutate("intervalCode"="A") |>
+  rename("stiffnessROC"="stiffness_ROC_A") |>
+  rename("stiffnessFrac"="stiffness_frac_A") |>
+  rename("dampingROC"="damping_ROC_A") |>
+  rename("dampingFrac"="damping_frac_A")
+foamDataB <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_B, E1_MPA_Cyc_B, 
+                    stiffness_ROC_B, stiffness_frac_B, delta_rad_Cyc_B, damping_ROC_B, 
+                    damping_frac_B, NVP_Cyc_B) |>
+  rename("cycle"="Cyc_B") |>
+  rename("stiffness"="E1_MPA_Cyc_B") |>
+  rename("damping"="delta_rad_Cyc_B") |>
+  rename("NVP"="NVP_Cyc_B") |>
+  mutate("intervalCode"="B") |>
+  rename("stiffnessROC"="stiffness_ROC_B") |>
+  rename("stiffnessFrac"="stiffness_frac_B") |>
+  rename("dampingROC"="damping_ROC_B") |>
+  rename("dampingFrac"="damping_frac_B")
+foamDataC <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_C, E1_MPA_Cyc_C, 
+                    stiffness_ROC_C, stiffness_frac_C, delta_rad_Cyc_C, damping_ROC_C, 
+                    damping_frac_C, NVP_Cyc_C) |>
+  rename("cycle"="Cyc_C") |>
+  rename("stiffness"="E1_MPA_Cyc_C") |>
+  rename("damping"="delta_rad_Cyc_C") |>
+  rename("NVP"="NVP_Cyc_C") |>
+  mutate("intervalCode"="C") |>
+  rename("stiffnessROC"="stiffness_ROC_C") |>
+  rename("stiffnessFrac"="stiffness_frac_C") |>
+  rename("dampingROC"="damping_ROC_C") |>
+  rename("dampingFrac"="damping_frac_C")
+foamDataD <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_D, E1_MPA_Cyc_D, 
+                    stiffness_ROC_D, stiffness_frac_D, delta_rad_Cyc_D, damping_ROC_D, 
+                    damping_frac_D, NVP_Cyc_D) |>
+  rename("cycle"="Cyc_D") |>
+  rename("stiffness"="E1_MPA_Cyc_D") |>
+  rename("damping"="delta_rad_Cyc_D") |>
+  rename("NVP"="NVP_Cyc_D") |>
+  mutate("intervalCode"="D") |>
+  rename("stiffnessROC"="stiffness_ROC_D") |>
+  rename("stiffnessFrac"="stiffness_frac_D") |>
+  rename("dampingROC"="damping_ROC_D") |>
+  rename("dampingFrac"="damping_frac_D")
+foamDataE <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_E, E1_MPA_Cyc_E, 
+                    stiffness_ROC_E, stiffness_frac_E, delta_rad_Cyc_E, damping_ROC_E, 
+                    damping_frac_E, NVP_Cyc_E) |>
+  rename("cycle"="Cyc_E") |>
+  rename("stiffness"="E1_MPA_Cyc_E") |>
+  rename("damping"="delta_rad_Cyc_E") |>
+  rename("NVP"="NVP_Cyc_E") |>
+  mutate("intervalCode"="E") |>
+  rename("stiffnessROC"="stiffness_ROC_E") |>
+  rename("stiffnessFrac"="stiffness_frac_E") |>
+  rename("dampingROC"="damping_ROC_E") |>
+  rename("dampingFrac"="damping_frac_E")
+foamDataF <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_F, E1_MPA_Cyc_F, 
+                    stiffness_ROC_F, stiffness_frac_F, delta_rad_Cyc_F, damping_ROC_F, 
+                    damping_frac_F, NVP_Cyc_F) |>
+  rename("cycle"="Cyc_F") |>
+  rename("stiffness"="E1_MPA_Cyc_F") |>
+  rename("damping"="delta_rad_Cyc_F") |>
+  rename("NVP"="NVP_Cyc_F") |>
+  mutate("intervalCode"="F") |>
+  rename("stiffnessROC"="stiffness_ROC_F") |>
+  rename("stiffnessFrac"="stiffness_frac_F") |>
+  rename("dampingROC"="damping_ROC_F") |>
+  rename("dampingFrac"="damping_frac_F")
+foamDataG <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_G, E1_MPA_Cyc_G, 
+                    stiffness_ROC_G, stiffness_frac_G, delta_rad_Cyc_G, damping_ROC_G, 
+                    damping_frac_G, NVP_Cyc_G) |>
+  rename("cycle"="Cyc_G") |>
+  rename("stiffness"="E1_MPA_Cyc_G") |>
+  rename("damping"="delta_rad_Cyc_G") |>
+  rename("NVP"="NVP_Cyc_G") |>
+  mutate("intervalCode"="G") |>
+  rename("stiffnessROC"="stiffness_ROC_G") |>
+  rename("stiffnessFrac"="stiffness_frac_G") |>
+  rename("dampingROC"="damping_ROC_G") |>
+  rename("dampingFrac"="damping_frac_G")
+foamDataH <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_H, E1_MPA_Cyc_H, 
+                    stiffness_ROC_H, stiffness_frac_H, delta_rad_Cyc_H, damping_ROC_H, 
+                    damping_frac_H, NVP_Cyc_H) |>
+  rename("cycle"="Cyc_H") |>
+  rename("stiffness"="E1_MPA_Cyc_H") |>
+  rename("damping"="delta_rad_Cyc_H") |>
+  rename("NVP"="NVP_Cyc_H") |>
+  mutate("intervalCode"="H") |>
+  rename("stiffnessROC"="stiffness_ROC_H") |>
+  rename("stiffnessFrac"="stiffness_frac_H") |>
+  rename("dampingROC"="damping_ROC_H") |>
+  rename("dampingFrac"="damping_frac_H")
+foamDataI <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_I, E1_MPA_Cyc_I, 
+                    stiffness_ROC_I, stiffness_frac_I, delta_rad_Cyc_I, damping_ROC_I, 
+                    damping_frac_I, NVP_Cyc_I) |>
+  rename("cycle"="Cyc_I") |>
+  rename("stiffness"="E1_MPA_Cyc_I") |>
+  rename("damping"="delta_rad_Cyc_I") |>
+  rename("NVP"="NVP_Cyc_I") |>
+  mutate("intervalCode"="I") |>
+  rename("stiffnessROC"="stiffness_ROC_I") |>
+  rename("stiffnessFrac"="stiffness_frac_I") |>
+  rename("dampingROC"="damping_ROC_I") |>
+  rename("dampingFrac"="damping_frac_I")
+foamDataJ <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_J, E1_MPA_Cyc_J, 
+                    stiffness_ROC_J, stiffness_frac_J, delta_rad_Cyc_J, damping_ROC_J, 
+                    damping_frac_J, NVP_Cyc_J) |>
+  rename("cycle"="Cyc_J") |>
+  rename("stiffness"="E1_MPA_Cyc_J") |>
+  rename("damping"="delta_rad_Cyc_J") |>
+  rename("NVP"="NVP_Cyc_J") |>
+  mutate("intervalCode"="J") |>
+  rename("stiffnessROC"="stiffness_ROC_J") |>
+  rename("stiffnessFrac"="stiffness_frac_J") |>
+  rename("dampingROC"="damping_ROC_J") |>
+  rename("dampingFrac"="damping_frac_J")
+foamDataK <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_K, E1_MPA_Cyc_K, 
+                    stiffness_ROC_K, stiffness_frac_K, delta_rad_Cyc_K, damping_ROC_K, 
+                    damping_frac_K, NVP_Cyc_K) |>
+  rename("cycle"="Cyc_K") |>
+  rename("stiffness"="E1_MPA_Cyc_K") |>
+  rename("damping"="delta_rad_Cyc_K") |>
+  rename("NVP"="NVP_Cyc_K") |>
+  mutate("intervalCode"="K") |>
+  rename("stiffnessROC"="stiffness_ROC_K") |>
+  rename("stiffnessFrac"="stiffness_frac_K") |>
+  rename("dampingROC"="damping_ROC_K") |>
+  rename("dampingFrac"="damping_frac_K")
+foamDataL <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_L, E1_MPA_Cyc_L, 
+                    stiffness_ROC_L, stiffness_frac_L, delta_rad_Cyc_L, damping_ROC_L, 
+                    damping_frac_L, NVP_Cyc_L) |>
+  rename("cycle"="Cyc_L") |>
+  rename("stiffness"="E1_MPA_Cyc_L") |>
+  rename("damping"="delta_rad_Cyc_L") |>
+  rename("NVP"="NVP_Cyc_L") |>
+  mutate("intervalCode"="L") |>
+  rename("stiffnessROC"="stiffness_ROC_L") |>
+  rename("stiffnessFrac"="stiffness_frac_L") |>
+  rename("dampingROC"="damping_ROC_L") |>
+  rename("dampingFrac"="damping_frac_L")
+foamDataM <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_M, E1_MPA_Cyc_M, 
+                    stiffness_ROC_M, stiffness_frac_M, delta_rad_Cyc_M, damping_ROC_M, 
+                    damping_frac_M, NVP_Cyc_M) |>
+  rename("cycle"="Cyc_M") |>
+  rename("stiffness"="E1_MPA_Cyc_M") |>
+  rename("damping"="delta_rad_Cyc_M") |>
+  rename("NVP"="NVP_Cyc_M") |>
+  mutate("intervalCode"="M") |>
+  rename("stiffnessROC"="stiffness_ROC_M") |>
+  rename("stiffnessFrac"="stiffness_frac_M") |>
+  rename("dampingROC"="damping_ROC_M") |>
+  rename("dampingFrac"="damping_frac_M")
+foamDataN <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_N, E1_MPA_Cyc_N, 
+                    stiffness_ROC_N, stiffness_frac_N, delta_rad_Cyc_N, damping_ROC_N, 
+                    damping_frac_N, NVP_Cyc_N) |>
+  rename("cycle"="Cyc_N") |>
+  rename("stiffness"="E1_MPA_Cyc_N") |>
+  rename("damping"="delta_rad_Cyc_N") |>
+  rename("NVP"="NVP_Cyc_N") |>
+  mutate("intervalCode"="N") |>
+  rename("stiffnessROC"="stiffness_ROC_N") |>
+  rename("stiffnessFrac"="stiffness_frac_N") |>
+  rename("dampingROC"="damping_ROC_N") |>
+  rename("dampingFrac"="damping_frac_N")
+
+foamData <- rbind(foamDataA, foamDataB, foamDataC, foamDataD, foamDataE, foamDataF, foamDataG, 
+                  foamDataH, foamDataI, foamDataJ, foamDataK, foamDataL, foamDataM, foamDataN)
+
+# Filter rows with NA in the cycle, stiffness, and damping columns
+foamData <- filter(foamData, !is.na(cycle))
+
+# Reorder the columns
+foamData <- select(foamData, 3, 1, 2, 4, 5, 6, 7, 16, 8, 11, 14, 9, 12, 10, 13, 15)
+
+# write the file
+write_csv(foamData, "foamData.csv")
+```
+
+### R Script for plots, models, and tables in Sections 2 and 3:
+
+```{r}
+#| echo: true
+#| eval: false
+
 foam_data <- read.csv("foamData.csv")
 
+#Some extra wrangling
 foam_data_scaled <- foam_data |>
@@
 
-FactorModel <- lmer(log(stiffnessFrac) ~  porosity + factor(chemIndex) + factor(chemIndex):porosity +  log(cycle) +  factor(chemIndex):log(cycle) + porosity:log(cycle) + factor(chemIndex):porosity:log(cycle) + (1 + log(cycle)|ID), data = foam_data_scaled)
+FactorModel <- lmer(log(stiffnessFrac) ~  porosity + factor(chemIndex) 
+                    + factor(chemIndex):porosity + log(cycle) +  factor(chemIndex):log(cycle) 
+                    + porosity:log(cycle) + factor(chemIndex):porosity:log(cycle) 
+                    + (1 + log(cycle)|ID), data = foam_data_scaled)
 
-dampingModel <- lmer(log(dampingFrac) ~ porosity + factor(chemIndex) + freq + log(cycle) + porosity:log(cycle) + freq:log(cycle) + factor(chemIndex):log(cycle) + (1 + log(cycle)|ID), data = foam_data_scaled)
+dampingModel <- lmer(log(dampingFrac) ~ porosity + factor(chemIndex) + freq + log(cycle) 
+                     + porosity:log(cycle) + freq:log(cycle) + factor(chemIndex):log(cycle) 
+                     + (1 + log(cycle)|ID), data = foam_data_scaled)
 
 # Convert intervalCode to a factor with specified levels
-foam_data_scaled$intervalCode <- factor(foam_data_scaled$intervalCode, levels = c("A", "B", "C", "D","E", "F", "G", "H","I", "J", "K", "L","M", "N"))
+foam_data_scaled$intervalCode <- factor(foam_data_scaled$intervalCode, levels = c("A", "B", "C", 
+                                       "D","E", "F", "G", "H","I", "J", "K", "L","M", "N"))
 
@@
 # graph1
-foam_Samples <- foam_data[ ,c("ID", "chemIndex", "porosity", "freq", "amp", "initialStrain", "initialStress")]
+foam_Samples <- foam_data[ ,c("ID", "chemIndex", "porosity", "freq", "amp", "initialStrain", 
+                              "initialStress")]
 foam_Samples <- unique(foam_Samples)
+
+count(foam_Samples, chemIndex, porosity)
 
@@
 # graph3
-sp1 <- ggplot(foam_data_scaled, aes(y = log(stiffnessFrac), x = log(cycle))) + geom_point(aes(color = freq))
-#sp2 <- ggplot(foam_data_scaled, aes(y = log(stiffnessFrac), x = log(cycle))) + geom_point(aes(color = factor(chemIndex)))
-#sp3 <- ggplot(foam_data_scaled, aes(y = log(stiffnessFrac), x = log(cycle))) + geom_point(aes(color = porosity))
+sp1 <- ggplot(foam_data_scaled, aes(y = log(stiffnessFrac), x = log(cycle))) 
++ geom_point(aes(color = freq))
 
-#combined <- sp1 + sp2 + sp3
-
-#combined
-
-dp1 <- ggplot(foam_data_scaled, aes(y = log(dampingFrac), x = log(cycle))) + geom_point(aes(color = freq))
-#dp2 <- ggplot(foam_data_scaled, aes(y = log(dampingFrac), x = log(cycle))) + geom_point(aes(color = factor(chemIndex)))
-#dp3 <- ggplot(foam_data_scaled, aes(y = log(dampingFrac), x = log(cycle))) + geom_point(aes(color = porosity))
-
-#combined <- dp1 + dp2 + dp3
-
-#combined
+dp1 <- ggplot(foam_data_scaled, aes(y = log(dampingFrac), x = log(cycle))) 
++ geom_point(aes(color = freq))
 
@@
 # graph4
-sfc1 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 1 & foam_data$porosity == "high",], aes(x=log(cycle), y=log(stiffnessFrac))) +
+sfc1 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 1 & foam_data$porosity 
+                                == "high",], aes(x=log(cycle), y=log(stiffnessFrac))) +
   geom_point(color="red", size=3) + 
@@
   theme(plot.title = element_text(size=25))
-sfc2 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 2 & foam_data$porosity == "high",], aes(x=log(cycle), y=log(stiffnessFrac))) +
+sfc2 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 2 & foam_data$porosity 
+                                == "high",], aes(x=log(cycle), y=log(stiffnessFrac))) +
   geom_point(color="green", size=3) + 
@@
   theme(plot.title = element_text(size=25))
-sfc3 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 3 & foam_data$porosity == "high",], aes(x=log(cycle), y=log(stiffnessFrac))) +
+sfc3 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 3 & foam_data$porosity 
+                                == "high",], aes(x=log(cycle), y=log(stiffnessFrac))) +
   geom_point(color="cyan2", size=3) + 
@@
   theme(plot.title = element_text(size=25))
-sfc4 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 1 & foam_data$porosity == "low",], aes(x=log(cycle), y=log(stiffnessFrac))) +
+sfc4 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 1 & foam_data$porosity 
+                                == "low",], aes(x=log(cycle), y=log(stiffnessFrac))) +
   geom_point(color="red3", size=3) + 
@@
   theme(plot.title = element_text(size=25))
-sfc5 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 2 & foam_data$porosity == "low",], aes(x=log(cycle), y=log(stiffnessFrac))) +
+sfc5 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 2 & foam_data$porosity
+                                == "low",], aes(x=log(cycle), y=log(stiffnessFrac))) +
   geom_point(color="green4", size=3) + 
@@
   theme(plot.title = element_text(size=25))
-sfc6 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 3 & foam_data$porosity == "low",], aes(x=log(cycle), y=log(stiffnessFrac))) +
+sfc6 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 3 & foam_data$porosity
+                                == "low",], aes(x=log(cycle), y=log(stiffnessFrac))) +
   geom_point(col="blue", size=3) + 
@@
 # graph5
-dfc1 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 1 & foam_data_scaled$porosity == "high",], aes(x=log(cycle), y=log(dampingFrac))) +
+dfc1 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 1 & foam_data_scaled$porosity 
+                                == "high",], aes(x=log(cycle), y=log(dampingFrac))) +
   geom_point(color="red", size=3) + 
@@
   theme(plot.title = element_text(size=25))
-dfc2 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 2 & foam_data_scaled$porosity == "high",], aes(x=log(cycle), y=log(dampingFrac))) +
+dfc2 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 2 & foam_data_scaled$porosity 
+                                == "high",], aes(x=log(cycle), y=log(dampingFrac))) +
   geom_point(color="green", size=3) + 
@@
   theme(plot.title = element_text(size=25))
-dfc3 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 3 & foam_data_scaled$porosity == "high",], aes(x=log(cycle), y=log(dampingFrac))) +
+dfc3 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 3 & foam_data_scaled$porosity 
+                                == "high",], aes(x=log(cycle), y=log(dampingFrac))) +
   geom_point(color="cyan2", size=3) + 
@@
   theme(plot.title = element_text(size=25))
-dfc4 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 1 & foam_data_scaled$porosity == "low",], aes(x=log(cycle), y=log(dampingFrac))) +
+dfc4 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 1 & foam_data_scaled$porosity 
+                                == "low",], aes(x=log(cycle), y=log(dampingFrac))) +
   geom_point(color="red3", size=3) + 
@@
   theme(plot.title = element_text(size=25))
-dfc5 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 2 & foam_data_scaled$porosity == "low",], aes(x=log(cycle), y=log(dampingFrac))) +
+dfc5 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 2 & foam_data_scaled$porosity 
+                                == "low",], aes(x=log(cycle), y=log(dampingFrac))) +
   geom_point(color="green4", size=3) + 
@@
   theme(plot.title = element_text(size=25))
-dfc6 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 3 & foam_data_scaled$porosity == "low",], aes(x=log(cycle), y=log(dampingFrac))) +
+dfc6 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 3 & foam_data_scaled$porosity 
+                                == "low",], aes(x=log(cycle), y=log(dampingFrac))) +
   geom_point(col="blue", size=3) + 
@@
 
-# summary of stiffnessFrac model 
-summary(FactorModel)
-
-# residuals and QQ plot of stiffnessFrac model
-plot(FactorModel)
-qqnorm(resid(FactorModel))
-qqline(resid(FactorModel))
+Anova(FactorModel)
 
@@
 
-# summary of dampingFrac model
-summary(dampingModel)
-
-# residuals and QQ plot of dampingFrac model
-plot(dampingModel)
-qqnorm(resid(dampingModel))
-qqline(resid(dampingModel))
+Anova(dampingModel)
 
@@
 
-```{r, include=TRUE}
-foam_Samples <- foam_data[ ,c("ID", "chemIndex", "porosity", "freq", "amp", "initialStrain", "initialStress")]
-foam_Samples <- unique(foam_Samples)
-
-count(foam_Samples, chemIndex)
-count(foam_Samples, porosity)
-count(foam_Samples, chemIndex, porosity)
-```
-
 ```{r}
-#| echo: false
+#| echo: true
 #| label: fig-example1
@@
 #| fig-width: 7
-#| fig-height: 5
-regfit.full <- regsubsets(log(stiffnessFrac) ~ chemIndex + porosity + freq + amp + initialStrain + initialStress + log(cycle) + (chemIndex + porosity + freq + amp + initialStrain + initialStress)*log(cycle) + (1 + log(cycle)|ID), data = foam_data_scaled)
+#| fig-height: 4
+regfit.full <- regsubsets(log(stiffnessFrac) ~ chemIndex + porosity + freq + amp 
+                          + initialStrain + initialStress + log(cycle) + (chemIndex 
+                          + porosity + freq + amp + initialStrain + initialStress)*log(cycle) 
+                          + (1 + log(cycle)|ID), 
+                          data = foam_data_scaled)
 #regfit.full <- regsubsets(stiffnessFrac ~ ., data = foam_data_scaled)
@@
 ```{r}
-#| echo: false
+#| echo: true
 #| label: fig-example2
@@
 #| fig-width: 7
-#| fig-height: 5
+#| fig-height: 4
 #| 
-regfit.full <- regsubsets(log(dampingFrac) ~ chemIndex + porosity + freq + amp + initialStrain + initialStress + log(cycle) + chemIndex:porosity + (chemIndex + porosity + freq + amp + initialStrain + initialStress + chemIndex:porosity)*log(cycle) + (1 + log(cycle)|ID), data = foam_data_scaled)
+regfit.full <- regsubsets(log(dampingFrac) ~ chemIndex + porosity + freq + amp 
+                          + initialStrain + initialStress + log(cycle) + chemIndex:porosity 
+                          + (chemIndex + porosity + freq + amp + initialStrain 
+                          + initialStress + chemIndex:porosity)*log(cycle) 
+                          + (1 + log(cycle)|ID), data = foam_data_scaled)
 
@@
 
-```{r, echo = FALSE}
-#| echo: false
+```{r}
+#| echo: true
 #| label: fig-example3
@@
 
-```{r, echo = FALSE}
-#| echo: false
+```{r}
+#| echo: true
 #| label: fig-example4
@@
 
-```{r, echo = FALSE}
-#| echo: false
+```{r}
+#| echo: true
 #| label: fig-example5
@@
 
-```{r, echo = FALSE}
-#| echo: false
+```{r}
+#| echo: true
 #| label: fig-example6

### 2024-05-07 CS2 Final Draft.qmd
Update CS2 Final Draft.qmd

@@
 
-In order to select our variables, we first started with a mixed model where all variables aside from log(cycle) were treated as level two variables and log(cycle) was treated as a level one variable.
-In this context, the level two variables on their own will impact the intercept of our model and the interaction between log(cycle) and the other variables will impact the slope of our model.
+To select our variables, we started with a mixed model where all variables aside from log(cycle) were treated as level two variables and log(cycle) was treated as a level one variable.
+The level two variables on their own will impact the intercept of our model and the interaction between log(cycle) and the other variables will impact the slope of our model.
 After starting with this full model, we manually removed variables from the model to increase the significance of the coefficients and to minimize the REML.
-From this process, we were able to create a model using the following predictors:
+Once a base model was created, we then used ANOVA testing to determine if the different levels of each predictor were significant and to compare against the full model.
+Based on this testing, we were able to create a model using the following predictors:
 
@@
 
-\begin{center}
-\text{Table 4: Predictors Selected For Each Model}
-\end{center}
+For the purpose of this paper, these are the variables we will be using in our model.
+We did find that there is a significant relationship between stiffness fraction and an interaction between amplitude and frequency.
+Unfortunately, the addition of this term lowered the significance of the relationship between the response and chemical index.
+Since chemical index is one of the focuses of this experiment, we decided to continue with our original model instead of including this term.
 
@@
 
-**Level 1**: $\log(\text{stiffnessFrac}) = \alpha_i + \beta_i*log(cycle)$
-
-**Level 2**: $\alpha_i = \alpha_0 + \alpha_1*\text{chemIndex} + \alpha_2*\text{freq} + \alpha_3*\text{amp}$
-
+Level 1: $\log(\text{stiffnessFrac}) = \alpha_i + \beta_i*log(cycle)$
+Level 2: $\alpha_i = \alpha_0 + \alpha_1*\text{chemIndex} + \alpha_2*\text{freq} + \alpha_3*\text{amp}$
          $\beta_i = \beta_0 + \beta_1*\text{chemIndex} +\beta_2*\text{freq} + \beta_3*\text{amp}$
 
-From the statistical summary below we see a high significance in our coefficients.
-The p-value for every term has a significance below five percent, meaning that there is a significant relationship between each of the terms and the response variable.
-The REML is also extremely low which leads us to conclude that this model fits the data very well and has minimal error.
-The coefficients not relating to cycle are impacting the initial stiffness of the foam at time zero.
-The variable that has the most positive impact on the initial stiffness is having a chemical index of 2.
+All samples have a stiffness fraction of 1 at cycle zero.
+The structure of a mixed model creates individual intercepts for each sample so for this model, we will be using samples starting at the first cycle measurement after cycle zero and going to cycle 10000.
+The alpha terms in the model will explain the impact of the covariates at the initial time recorded and the beta terms will explain the impact of the covariates on the stiffness fraction as time goes on.
+
+From @Table-1 we see a high significance in our coefficients.
+Although not every variable is significant on it's own, the interaction with time is.
+For example, chemical index does not appear to have a significant relationship with the response on it's own, but one level of chemical index has a significant relationship when interacted with cycle.
+The interaction terms in this model are an indicator of how the response will change based on both a covariate and time.
+From this summary, amplitude appears to have the largest positive impact on initial stiffness fraction and the largest negative impact as time progresses.
 
 ```{r, echo = FALSE}
-Anova(FactorModel)
-```
-
-We also assessed the model to see how it met the conditions of a predictive model.
-From @fig-example3 (see Appendix), we see that there appears to be some sinusoidal behavior at the beginning and the middle of the shape has a larger range than the rest of the plot.
-Since the range of the plot is so small, we can conclude that the equal variance condition is being met.
-We also do not see any flaring or nonlinear movement so the linearity condition is also being met.
-The QQ plot in @fig-example4 (see Appendix) shows some tailing on the ends but nothing too concerning.
-This concludes that all conditions are being met so this model is valid for the purpose of prediction.
+#| echo: false
+#| label: Table-1
+#| warning: false
+#| fig-cap: "Coefficients for Stiffness Fraction Model"
+#| fig-align: "center"
+#| fig-width: 7
+#| fig-height: 4
+library(car)
+suppressWarnings(FactorModel1 <- lmer(log(stiffnessFrac) ~ factor(chemIndex) + freq + amp +  log(cycle) +  factor(chemIndex):log(cycle) +  freq:log(cycle) + amp:log(cycle) + (1 + log(cycle)|ID), data = foam_data_scaled))
+summary(FactorModel1)$coef
+```
+
+Although @Table-1 indicates there is significance to the different levels of chemical index, the output below in @Table-2 confirms that there is a significant difference between at least on of the levels of chemical index.
+To further assess the significane of the model, we explored the residual plots and found that this model meets the linearity, normality, and equal variance conditions.
+These plots can be further explored in the appendix.
+
+```{r, echo = FALSE}
+#| echo: false
+#| label: Table-2
+#| warning: false
+#| fig-cap: "Anova Analysis of Stiffness Fraction Model"
+#| fig-align: "center"
+#| fig-width: 7
+#| fig-height: 4
+Anova(FactorModel1)
+```
 
@@
 
-**Question 1:** We see that chemical index, frequency, and cycle number have the most significant effects on both damping and stiffness.
-However porosity has an effect on damping and amplitude as an effect on stiffness.
+**Question 1:** We see that frequency, amplitude, chemical index, and cycle all have significant impacts on both stiffness and damping.
+Porosity also has a significant impact on damping, but not on stiffness.
 
 **Question 2:** For stiffness, we see that the individual predictors on their own will increase the initial stiffness, but as time goes on, foam using chemical index 121 will have the greatest decrease in stiffness.
-The individual predictors will impact damping in a similar way but foam with a chemical index of 79 will see the greatest decay in damping.
+The individual predictors will impact damping in a similar way but foam with a chemical index of 79 and a low porosity will see the greatest decay in damping.
 Overall, there appear to be significant impacts on the stiffness and damping values depending on chemical index and porosity.
@@
 
+#StiffnessFrac Model
+library(car)
+suppressWarnings(FactorModel1 <- lmer(log(stiffnessFrac) ~ factor(chemIndex) + freq + amp +  log(cycle) +  factor(chemIndex):log(cycle) +  freq:log(cycle) + amp:log(cycle) + (1 + log(cycle)|ID), data = foam_data_scaled))
+summary(FactorModel1)$coef
 Anova(FactorModel)

### 2024-05-07 CS2 Final Draft.qmd
Update CS2 Final Draft.qmd
Proofreading

@@
 
-#  2. Exploratory Data Analysis (EDA)
+##  2. Exploratory Data Analysis (EDA)
 
@@
 
-# 3. Statistical Analysis
-
-Based on the initial research questions, we will be developing two separate models: one for stiffness fraction and the other for damping fraction.
-Due to the experimental design and the fact that our observations are not independent of each other, we used a mixed model format for our predictions.
-
-To select our variables, we started with a mixed model where all variables aside from log(cycle) were treated as level two variables and log(cycle) was treated as a level one variable.
-The level two variables on their own will impact the intercept of our model and the interaction between log(cycle) and the other variables will impact the slope of our model.
-After starting with this full model, we manually removed variables from the model to increase the significance of the coefficients and to minimize the REML.
-Once a base model was created, we then used ANOVA testing to determine if the different levels of each predictor were significant and to compare against the full model.
-Based on this testing, we were able to create a model using the following predictors:
+## 3. Statistical Analysis
+
+Based on the research questions, we developed two separate models: one for stiffness fraction and the other for damping fraction.
+Due to the experimental design, and the fact that our observations are not independent of each other, we used a mixed model format for our predictions.
+Log(cycle) was treated as a level-one variable because it changes across individual measurements, while all other variables were treated as level-two variables because they only vary across foam samples.
+In a mixed model, level-two variables impact the model intercept, and level-one variables interact with the level-two variables to effect the model slope.
+
+After starting with full models, we manually removed variables to increase the significance of the coefficients and to minimize the REML.
+Once base models were created, we used ANOVA testing to determine if the different levels of each predictor were significant and to compare against the full models.
+Based on this testing, we created models with the following predictors:
 
@@
 
-For the purpose of this paper, these are the variables we will be using in our model.
-We did find that there is a significant relationship between stiffness fraction and an interaction between amplitude and frequency.
-Unfortunately, the addition of this term lowered the significance of the relationship between the response and chemical index.
-Since chemical index is one of the focuses of this experiment, we decided to continue with our original model instead of including this term.
-
-# 3.1 Stiffness
-
-To best predict the value of log(stiffness fraction), we used the following model:
-
-Level 1: $\log(\text{stiffnessFrac}) = \alpha_i + \beta_i*log(cycle)$
-Level 2: $\alpha_i = \alpha_0 + \alpha_1*\text{chemIndex} + \alpha_2*\text{freq} + \alpha_3*\text{amp}$
+We also found a significant relationship between stiffness fraction and the interaction between amplitude and frequency, but the addition of this term lowered the significance of the relationship between the response and chemical index.
+Since understanding the impact of chemical index is important to Foster, we decided to continue with our original model instead of including this term.
+
+## 3.1 Stiffness
+
+To predict log(stiffnessFrac), we used the following model:
+
+**Level 1**: $\log(\text{stiffnessFrac}) = \alpha_i + \beta_i*log(cycle)$
+
+**Level 2**: $\alpha_i = \alpha_0 + \alpha_1*\text{chemIndex} + \alpha_2*\text{freq} + \alpha_3*\text{amp}$
+
 $\beta_i = \beta_0 + \beta_1*\text{chemIndex} +\beta_2*\text{freq} + \beta_3*\text{amp}$
@@
 All samples have a stiffness fraction of 1 at cycle zero.
-The structure of a mixed model creates individual intercepts for each sample so for this model, we will be using samples starting at the first cycle measurement after cycle zero and going to cycle 10000.
-The alpha terms in the model will explain the impact of the covariates at the initial time recorded and the beta terms will explain the impact of the covariates on the stiffness fraction as time goes on.
-
-From @Table-1 we see a high significance in our coefficients.
-Although not every variable is significant on it's own, the interaction with time is.
-For example, chemical index does not appear to have a significant relationship with the response on it's own, but one level of chemical index has a significant relationship when interacted with cycle.
-The interaction terms in this model are an indicator of how the response will change based on both a covariate and time.
-From this summary, amplitude appears to have the largest positive impact on initial stiffness fraction and the largest negative impact as time progresses.
+A mixed model can create individual intercepts for each sample, so for this model, we used samples starting at the first cycle measurement after cycle zero.
+Thus, $\alpha$ terms in the model  explain the impact of the covariates right after compressions begin, and the $\beta$ terms explain the impact of the covariates as time goes on.
+
+From @Table-1, we see a high significance in our coefficients.
+Some variables are not significant on their own, their interactions with log(cycle) are.
+For example, chemical index does not appear to have a significant relationship with the response on it's own, but one level of chemical index has a significant relationship when interacted with log(cycle).
+The interaction terms explain how log(cycle) effects stiffness fraction differently under different conditions.
+From this summary, amplitude appears to have the largest positive impact on initial stiffness fraction, and the largest negative impact as time progresses.
 
@@
 
-Although @Table-1 indicates there is significance to the different levels of chemical index, the output below in @Table-2 confirms that there is a significant difference between at least on of the levels of chemical index.
-To further assess the significane of the model, we explored the residual plots and found that this model meets the linearity, normality, and equal variance conditions.
-These plots can be further explored in the appendix.
+Although @Table-1 indicates there is significance to the different levels of chemical index, the output below in @Table-2 confirms that there is a significant difference between at least one of the levels of chemical index.
+To further assess the significance of the model, we used diagnostic plots to find that this model meets the linearity, normality, and equal variance conditions.
+These plots are in the appendix.
 
@@
 
-Lastly, given that we are using a mixed effects model, it is necessary to explore the random effects associated with the model.
+Lastly, it was necessary to explore the random effects associated with the mixed model.
 From @fig-6, we notice that as the intercept increases, the slope decreases.
-This is a very defined trend and supports our conclusion that a mixed model should be used.
-We would expect to see that each of the foams starts at a similar stiffness fraction and as time increases, there would be deviations based on the other predictors.
+This is a very defined trend, supporting our conclusion that a mixed model should be used.
+We would expect the foams to start at a similar stiffness fraction, and deviate based on the other predictors as time passes.
 This plot confirms this behavior.
@@
 
-#	3.2 Damping
-
-To best predict the value of log(damping fraction), we used this model:
+##	3.2 Damping
+
+To predict log(damping fraction), we used this model:
 
@@
 
-We used a residuals vs fitted plot and QQ plot to assess the conditions of our model (see @fig-example5 and @fig-example6, in the Appendix).
+We used a residuals vs fitted plot and QQ plot to assess the conditions of our model (see Appendix).
 We found that this damping model meets all of these conditions which therefore we can analyze our model's results.
@@
 
-# 4. Recommendations
+## 4. Recommendations
 
@@
 
-# 5. Additional Considerations
+## 5. Additional Considerations
 
@@
 
-# Technical Appendix
+## Technical Appendix
 

### 2024-05-07 CS2 Final Draft.qmd
Update CS2 Final Draft.qmd

@@
 **Question 2:** How are these parameters impacting the fatigue response in stiffness and damping over time?
+
+\newpage
 
@@
 #| fig-width: 10
-#| fig-height: 5
+#| fig-height: 4
 
 sp1 <- ggplot(foam_data_scaled, aes(y = log(stiffnessFrac), x = log(cycle))) + geom_point(aes(color = freq))
-#sp2 <- ggplot(foam_data_scaled, aes(y = log(stiffnessFrac), x = log(cycle))) + geom_point(aes(color = factor(chemIndex)))
-#sp3 <- ggplot(foam_data_scaled, aes(y = log(stiffnessFrac), x = log(cycle))) + geom_point(aes(color = porosity))
-
-#combined <- sp1 + sp2 + sp3
-
-#combined
 
 dp1 <- ggplot(foam_data_scaled, aes(y = log(dampingFrac), x = log(cycle))) + geom_point(aes(color = freq))
-#dp2 <- ggplot(foam_data_scaled, aes(y = log(dampingFrac), x = log(cycle))) + geom_point(aes(color = factor(chemIndex)))
-#dp3 <- ggplot(foam_data_scaled, aes(y = log(dampingFrac), x = log(cycle))) + geom_point(aes(color = porosity))
-
-#combined <- dp1 + dp2 + dp3
-
-#combined
 
@@
 ```
+
+\newpage
 
@@
 
+\begin{center}
+\text{Table 4: Selected Predictors}
+\end{center}
+
 We also found a significant relationship between stiffness fraction and the interaction between amplitude and frequency, but the addition of this term lowered the significance of the relationship between the response and chemical index.
@@
 
-From @Table-1, we see a high significance in our coefficients.
+```{r, echo = FALSE}
+library(car)
+suppressWarnings(FactorModel1 <- lmer(log(stiffnessFrac) ~ factor(chemIndex) + freq + amp +  log(cycle) +  factor(chemIndex):log(cycle) +  freq:log(cycle) + amp:log(cycle) + (1 + log(cycle)|ID), data = foam_data_scaled))
+summary(FactorModel1)$coef
+```
+
+\begin{center}
+\text{Table 5: Coefficients for Stiffness Fraction Model}
+\end{center}
+
+In Table 5, we see high significance (aka p-value < 0.05) for most of our coefficients.
 Some variables are not significant on their own, their interactions with log(cycle) are.
@@
 
+Although Table 5 indicates chemical index levels 2 and 3 (aka 100 and 121) are not significant, the output below in Table 6 confirms that at least one chemical index is significantly different from another.
+To further assess the model, we used diagnostic plots to find that it meets the linearity, normality, and equal variance conditions.
+These plots are in the appendix.
+
 ```{r, echo = FALSE}
-#| echo: false
-#| label: Table-1
-#| warning: false
-#| fig-cap: "Coefficients for Stiffness Fraction Model"
-#| fig-align: "center"
-#| fig-width: 7
-#| fig-height: 4
-library(car)
-suppressWarnings(FactorModel1 <- lmer(log(stiffnessFrac) ~ factor(chemIndex) + freq + amp +  log(cycle) +  factor(chemIndex):log(cycle) +  freq:log(cycle) + amp:log(cycle) + (1 + log(cycle)|ID), data = foam_data_scaled))
-summary(FactorModel1)$coef
-```
-
-Although @Table-1 indicates there is significance to the different levels of chemical index, the output below in @Table-2 confirms that there is a significant difference between at least one of the levels of chemical index.
-To further assess the significance of the model, we used diagnostic plots to find that this model meets the linearity, normality, and equal variance conditions.
-These plots are in the appendix.
-
-```{r, echo = FALSE}
-#| echo: false
-#| label: Table-2
-#| warning: false
-#| fig-cap: "Anova Analysis of Stiffness Fraction Model"
-#| fig-align: "center"
-#| fig-width: 7
-#| fig-height: 4
 Anova(FactorModel1)
 ```
+
+\begin{center}
+\text{Table 6: Anova Analysis of Stiffness Fraction Model}
+\end{center}
 

### 2024-05-07 CS2 Final Draft.qmd
Add interpretations

@@
 Measurements taken beyond 10,000 cycles were removed to maintain consistency across observations, and because Moira identified this as the most relevant time frame.
+The first measurements for stiffnessFrac and dampingFrac were removed as well, because they were all equal to 1 and did not convey useful information.
 
@@
 
-ggplot(foam_Samples, aes(x=freq)) +
+ggplot(foam_Samples, aes(x=amp)) +
   geom_histogram(bins=10)
@@
 
-The histogram in @fig-1 shows the distribution of compression frequencies among the foam samples revealing that certain compression frequencies were tested more frequently than others.
+The histogram in @fig-1 shows the distribution of compression amplitudes among the foam samples, revealing that certain amplitudes were tested much less frequently than others.
 These non-uniformities in sample distribution might impact the reliability of conclusions drawn from the analysis.
@@
 ```
+
 In order to see how stiffnessFrac and dampingFrac change over time for all foams, we plotted a scatter plot for each vs cycle.
@@
 For chemical index 79, low porosity foam trends downward more often than high porosity foam, and for chemical index 100, low porosity foam trends upward more often than high porosity foam.
-So not only does porosity affect log(stiffnessFrac), but it effects it different ways for different porosities.
-This suggests that a porosity:chemIndex interaction term would improve our model.
+So not only does porosity affect log(stiffnessFrac), but it effects it different ways for different chemical indices This suggests that a porosity:chemIndex interaction term might improve our model.
 
@@
 Trends in log(dampingFrac) for chemical index 121 are more similar to each other for low porosity than for high porosity, and trends in log(dampingFrac) for chemical index 79 are bit more similar to each other for high porosity than for low porosity.
-This suggests a porosity:chemIndex interaction term would improve our damping model, but this is still less clear than it was for stiffness.
+This suggests a porosity:chemIndex interaction term could improve our damping model, but this is still less clear than it was for stiffness.
 
@@
 Based on the research questions, we developed two separate models: one for stiffness fraction and the other for damping fraction.
-Due to the experimental design, and the fact that our observations are not independent of each other, we used a mixed model format for our predictions.
+Due to the experimental design, and the fact that measurements from the same sample are not independent of each other, we used a mixed model format for our predictions.
 Log(cycle) was treated as a level-one variable because it changes across individual measurements, while all other variables were treated as level-two variables because they only vary across foam samples.
 In a mixed model, level-two variables impact the model intercept, and level-one variables interact with the level-two variables to effect the model slope.
+This accounts for each sample appeared to react to log(cycle) differently.
 
@@
 
+Despite the expectations we got from figures 4 and 5, we could not determine that an interaction between chemical index and porosity was as significant as the predictors listed in Table 4, so it was not included.
+
 ## 3.1 Stiffness
@@
 
-All samples have a stiffness fraction of 1 at cycle zero.
-A mixed model can create individual intercepts for each sample, so for this model, we used samples starting at the first cycle measurement after cycle zero.
-Thus, $\alpha$ terms in the model  explain the impact of the covariates right after compressions begin, and the $\beta$ terms explain the impact of the covariates as time goes on.
+The $\alpha$ terms in the model explain the impact of the covariates right after compressions begin (intercepts), and the $\beta$ terms explain the impact of the covariates as time goes on (slopes).
 
@@
 The interaction terms explain how log(cycle) effects stiffness fraction differently under different conditions.
-From this summary, amplitude appears to have the largest positive impact on initial stiffness fraction, and the largest negative impact as time progresses.
+
+In this summary, log(cycle) has a positive relationship with stiffnessFrac, but that is only when frequency and amplitude are zero.
+The interactions of freq and amp with log(cycle) have negative coefficients, showing us that stiffness decreases more when compressions are faster and larger.
+Chemical index levels 2 and 3 (aka 100 and 121) interacted with log(cycle) also have negative coefficients, indicating that these foams lose stiffness faster than foam with index 79.
 
@@
 
-A similar concern comes from foam types and testing conditions that were used less often, such as high frequency.
-The reliability of our models is already limited, and for these cases they are even less reliable.
-Additionally, if one variety of foam was never tested at high frequency, then the model's claims about that foam at high frequency are based on nothing, and should not be trusted at all.
+A similar concern comes from foam types and testing conditions that were used less often, such as high amplitude The reliability of our models is already limited, and for these cases they are even less reliable.
+Additionally, if one variety of foam was never tested at high amplitude, then the model's claims about that foam at high frequency are based on nothing, and should not be trusted at all.
 So not all information this model gives on what effects stiffness and damping in foam is equally trustworthy.

### 2024-05-07 CS2 Final Draft.qmd
Update CS2 Final Draft.qmd

@@
 
-**Level 2**: $\alpha_i = \alpha_0 + \alpha_1*\text{porosity} + \alpha_2*\text{chemIndex} + \alpha_3*\text{freq}$
-
-$\beta_i = \beta_0 + \beta_1*\text{porosity} + \beta_2*\text{chemIndex} +\beta_3*\text{freq}$
+**Level 2**: $\alpha_i = \alpha_0 + \alpha_1*\text{chemIndex} + \alpha_2*\text{freq}$
+
+$\beta_i = \beta_0 + \beta_1*\text{chemIndex} +\beta_2*\text{freq}$
+
+```{r, echo = FALSE}
+suppressWarnings(dampingModel <- lmer(log(stiffnessFrac) ~ factor(chemIndex) + freq + log(cycle) +  factor(chemIndex):log(cycle) + freq:log(cycle) + (1 + log(cycle)|ID), data = foam_data_scaled))
+summary(dampingModel)$coef
+```
+
+\begin{center}
+\text{Table 6: Coefficients for Damaping Fraction Model}
+\end{center}
 
@@
 These small p-values suggest that these variables are significant predictors of our response variable.
-The only value that does not show a significant p-value is chemIndex, which we chose to keep because the interaction between chemIndex and log(cycle) is significant.
-This model has multiple interactions: porosity and log(cycle), frequency and log(cycle), and chemical index and log(cycle).
-These interactions show that if all other variables are held constant, there is a combined effect of these interactions on the response variable.
-
-```{r, echo = FALSE}
+Like the stiffness model, there are some predictors that do not have a significant p-value but their interaction terms are very significant.
+We can double check our predictors in the below table as well and see that all our predictors are significant.
+
+```{r}
 Anova(dampingModel)
@@
 **Question 1:** We see that frequency, amplitude, chemical index, and cycle all have significant impacts on both stiffness and damping.
-Porosity also has a significant impact on damping, but not on stiffness.
 
 **Question 2:** For stiffness, we see that the individual predictors on their own will increase the initial stiffness, but as time goes on, foam using chemical index 121 will have the greatest decrease in stiffness.
-The individual predictors will impact damping in a similar way but foam with a chemical index of 79 and a low porosity will see the greatest decay in damping.
-Overall, there appear to be significant impacts on the stiffness and damping values depending on chemical index and porosity.
+The individual predictors will impact damping in a similar way but foam with a chemical index of 79 will see the greatest decay in damping.
+Overall, there appear to be significant impacts on the stiffness and damping values depending on chemical index and frequency.
 

### 2024-05-07 CS2 Final Draft.qmd
Update CS2 Final Draft.qmd
Merged changes

--- 2024-05-07 CS2 Final Draft.qmd
@@
 Chemical index levels 2 and 3 (aka 100 and 121) interacted with log(cycle) also have negative coefficients, indicating that these foams lose stiffness faster than foam with index 79.
+However, these two chemical indices have higher stiffness right after compressions start, and foams with high frequency and amplitude do as well.
 
@@
 
+```{r, echo = FALSE}
+suppressWarnings(dampingModel <- lmer(log(stiffnessFrac) ~ factor(chemIndex) + freq + log(cycle) +  factor(chemIndex):log(cycle) + freq:log(cycle) + (1 + log(cycle)|ID), data = foam_data_scaled))
+summary(dampingModel)$coef
+```
+
+\begin{center}
+\text{Table 7: Coefficients for Damaping Fraction Model}
+\end{center}
+
 The summary of this model shows a high significance in almost all of our coefficients through small p-values (less than 0.05).
 These small p-values suggest that these variables are significant predictors of our response variable.
-The only value that does not show a significant p-value is chemIndex, which we chose to keep because the interaction between chemIndex and log(cycle) is significant.
-This model has multiple interactions: porosity and log(cycle), frequency and log(cycle), and chemical index and log(cycle).
-These interactions show that if all other variables are held constant, there is a combined effect of these interactions on the response variable.
-
-```{r, echo = FALSE}
+Like the stiffness model, there are some predictors that do not have a significant p-value but their interaction terms are very significant.
+We can double check our predictors in the below table as well and see that all our predictors are significant.
+
+```{r}
 Anova(dampingModel)
 ```
+
+\begin{center}
+\text{Table 8: Anova Analysis of Damping Fraction Model}
+\end{center}
 
@@
 
-**Question 2:** For stiffness, we see that the individual predictors on their own will increase the initial stiffness, but as time goes on, foam using chemical index 121 will have the greatest decrease in stiffness.
+**Question 2:** High frequency, high amplitude, and chemical indices 100 and 121 all increase the initial stiffness, but as time goes on, these also lose stiffness the fastest.
+Foam with chemical index 121 has the greatest decrease in stiffness.
 The individual predictors will impact damping in a similar way but foam with a chemical index of 79 and a low porosity will see the greatest decay in damping.

### 2024-05-07 CS2 Final Draft.qmd
Update CS2 Final Draft.qmd

--- 2024-05-07 CS2 Final Draft.qmd
@@
           log(cycle)               |            log(cycle)
-              amp                  |             porosity
-       chemIndex:log(cycle)        |         chemIndex:log(cycle)
-        freq:log(cycle)            |           freq:log(cycle)
-        amp:log(cycle)             |         porosity:log(cycle)
+              amp                  |         chemIndex:log(cycle)
+       chemIndex:log(cycle)        |          freq:log(cycle)
+        freq:log(cycle)            |           
+        amp:log(cycle)             |
 
@@
 
-$\beta_i = \beta_0 + \beta_1*\text{chemIndex} +\beta_2*\text{freq}$
+$\beta_i = \beta_0  + \beta_2*\text{chemIndex} +\beta_3*\text{freq}$
 
@@
 \begin{center}
-\text{Table 6: Coefficients for Damaping Fraction Model}
+\text{Table 7: Coefficients for Damaping Fraction Model}
 \end{center}
@@
 ```
+
+\begin{center}
+\text{Table 8: Anova Analysis of Damping Fraction Model}
+\end{center}
 

### 2024-05-07 CS2 Final Draft.qmd
Update CS2 Final Draft.qmd

@@
 author: 'Katie Greene, Ava Harman, Chris Seery, Ian Stonecypher'
-date: 'May, 2024'
+date: '2024-05-06'
 format: pdf
@@
 #| fig-width: 4
-#| fig-height: 4
+#| fig-height: 3
 
@@
 
-ggplot(foam_Samples, aes(x=amp)) +
+ggplot(foam_Samples, aes(x=freq)) +
   geom_histogram(bins=10)
@@
 
-The histogram in @fig-1 shows the distribution of compression amplitudes among the foam samples, revealing that certain amplitudes were tested much less frequently than others.
+The histogram in @fig-1 shows the distribution of compression frequencies among the foam samples, revealing that certain frequencies were tested much less often than others.
 These non-uniformities in sample distribution might impact the reliability of conclusions drawn from the analysis.
@@
 In a mixed model, level-two variables impact the model intercept, and level-one variables interact with the level-two variables to effect the model slope.
-This accounts for each sample appeared to react to log(cycle) differently.
+This accounts for how each sample appeared to react to log(cycle) differently.
 
@@
 
-$\beta_i = \beta_0  + \beta_2*\text{chemIndex} +\beta_3*\text{freq}$
+$\beta_i = \beta_0  + \beta_1*\text{chemIndex} +\beta_2*\text{freq}$
 
@@
 These small p-values suggest that these variables are significant predictors of our response variable.
-Like the stiffness model, there are some predictors that do not have a significant p-value but their interaction terms are very significant.
-We can double check our predictors in the below table as well and see that all our predictors are significant.
+
+Like in the stiffnessFrac model, the interaction between log(cycle) and frequency has a negative coefficient, so damping decreases more at higher compression frequencies.
+Chemical index levels 2 and 3 (aka 100 and 121) interacted with log(cycle) also have negative coefficients, indicating that these foams lose damping faster than foam with index 79.
+But like in the stiffnessFrac model, these two chemical indices have higher damping right after compressions start, and foams with high frequency do as well.
+
+Like the stiffness model, not all chemical indices look significant in Table 7, but we can double-check our predictors in the Anova in Table 8 and see that all our predictors are significant.
 
@@
 #| fig-align: "center"
-#| fig-width: 7
+#| fig-width: 8
 #| fig-height: 4
@@
 
+\newpage
+
 ## 4. Recommendations
 
-**Question 1:** We see that frequency, amplitude, chemical index, and cycle all have significant impacts on both stiffness and damping.
-
-**Question 2:** For stiffness, we see that the individual predictors on their own will increase the initial stiffness, but as time goes on, foam using chemical index 121 will have the greatest decrease in stiffness.
-The individual predictors will impact damping in a similar way but foam with a chemical index of 79 will see the greatest decay in damping.
-Overall, there appear to be significant impacts on the stiffness and damping values depending on chemical index and frequency.
+**Question 1:** We see that frequency, chemical index, and cycle all have significant impacts on both stiffness and damping. Amplitude also has an impact on stiffness.
+
+**Question 2:** For stiffness, we see that high frequency, high amplitude, and chemical indices 100 and 121 will increase the stiffness right after compressions start, but as time goes on, these will cause the more decrease in stiffness.
+Foam using chemical index 121 will have the greatest decrease in stiffness, and foam using chemical index 79 will have its stiffness decrease the least.
+For damping, high frequency and chemical indices 100 and 121 similarly lead to higher damping right after compressions start, and lead to the fastest decay in damping.
+The type of foam with the greatest decay in damping is chemical index 121, and foam using chemical index 79 will have its damping decrease the least.
 
@@
 
-A similar concern comes from foam types and testing conditions that were used less often, such as high amplitude The reliability of our models is already limited, and for these cases they are even less reliable.
+A similar concern comes from testing conditions that were used less often, such as high frequency.
+The reliability of our models is already limited, and for these cases they are even less reliable.
 Additionally, if one variety of foam was never tested at high amplitude, then the model's claims about that foam at high frequency are based on nothing, and should not be trusted at all.
@@
 summary(FactorModel1)$coef
+
 Anova(FactorModel)
@@
 re <- ranef(FactorModel)$ID
-
 plot(re)
+
+suppressWarnings(dampingModel <- lmer(log(stiffnessFrac) ~ factor(chemIndex) + freq + log(cycle) +  factor(chemIndex):log(cycle) + freq:log(cycle) + (1 + log(cycle)|ID), data = foam_data_scaled))
+summary(dampingModel)$coef
 
@@
 re <- ranef(dampingModel)$ID
-
 plot(re)
